---
output: 
  bookdown::pdf_document2
linkcolor: blue
bibliography: C:/Users/colej/Documents/Research projects/bib_ch.bib
#biblio-style: "apalike"
fontsize: 12pt
link-citations: yes
toc: false
csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
indent: true
abstract: "Two models of the relationship between election manipulation and protest can be found in the literature. One approach holds that more severe election fraud reveals an incumbent who is more vulnerable to protest; in this view, protest risk can result in cleaner elections. An alternative approach contends that election manipulation can be a sign of underlying regime strengths; under this framework, protest risk is less likely to deter manipulation. This article interrogates the assumptions underlying these models, and conducts a novel test of their implications. Using data on the occurrence and duration of protest waves in 659 authoritarian cases from 1990-2012, this paper shows that governments which manipulate elections more also defeat protests more quickly. This supports the ‘regime-strength’ model, and helps explain why electoral authoritarianism is a sticky equilibrium amid ongoing democratic backsliding. It suggests researchers should look beyond protest risk in models of electoral manipulation."
header-includes: 
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{multirow}
  - \usepackage{lscape}
  - \usepackage{pdflscape}
  - \usepackage{longtable}
  - \usepackage[normalem]{ulem}
  - \useunder{\uline}{\ul}{}

---

\doublespace

<!--
The December 2011 elections to the Russian State Duma delivered a significant setback for the ruling party. It claimed only forty-nine percent of the vote, and while it retained the majority in the legislature, it lost more than 75 seats from the previous convocation. This narrow 'victory' occurred despite widespread election-day manipulation, detected in many cases by domestic election observers. The election triggered one of the most substantial protest movements in post-Soviet Russian history. More than 100,000 people turned out in protest on two occasions in Moscow, with tens of thousands more taking to the streets in regional cities [@Dollbaum2017]. Protests continued regularly in advance of the March 2012 presidential election, declining afterward.

Despite the significant challenge, the ruling party was able to defuse the protest movement over a period of months. Through a combination of repression and concessions [@golosov2012], the ruling party was able to reassert itself and continue a long process of sidelining the political opposition in Russia [@gelman_2015_troubled]. The post-2011 protest wave is one much-studied case of the phenomenon of electoral protest aimed at punishing electoral manipulation. Yet the Russian example highlights questions that get to the heart of a debate in the study of electoral manipulation and protest. When opposition movements protest manipulated election results, are they doing so because the incumbent\textemdash by relying on manufactured votes\textemdash has signaled its own weakness? This view is adopted by many researchers, especially in the formal modelling tradition. Or, does election manipulation indicate sources of strength for incumbents (independent of genuine popularity), and does protest occur for non-strategic reasons? 
!-->

It can be a stirring sight: following an unfair election, thousands of people take to the streets to demand that new, free elections. In some cases, the scale of the protests and solidarity of their participants can precipitate the collapse of a non-democratic regime; the 2004 Orange Revolution succeeded in preventing the theft of a presidential election in Ukraine, for instance, while large protests in Peru helped prevent President Alberto Fujimori from claiming a third term. In other cases, however---despite protesters' determination---incumbents are able to remain in power through a combination of repression and concessions. Large, courageous, but ultimately failed protest movements in Russia in 2011 and Belarus in 2020 offer stark examples.

Successful electoral revolutions figure prominently in theories of election manipulation and how it might be restrained, with broader implications for the stability of both democratic and authoritarian regimes. A classic view, influential in the formal modeling tradition, argues that the use of election manipulation reveals an incumbent's unpopularity, signaling weakness against a post-election challenge. As more extensive manipulation signals deeper weakness, strategic opposition actors should be more likely to initiate protest as election manipulation increases (all else equal). Such models predict that some incumbents, who require manipulation to win an election with high probability, choose instead to gamble on relatively clean elections out of fear of costly protest. This framework has normatively optimistic implications, by showing that stronger opposition movements\textemdash those that are more organized, better resourced, and more popular\textemdash will be better able to deter incumbents' from holding rigged elections. In turn, this dynamic makes democratic transition and consolidation more likely. Because fraud indicates incumbent weakness in this model, I refer to it as the regime-weakness model of electoral manipulation.

An alternative conceptualization of electoral manipulation has developed in recent years. In this view, developed most prominently by Simpser [-@simpser2013], election manipulation can be a costly signal of the incumbent's resources and capacities. When incumbents manipulate elections, they demonstrate financial, patronage, and organizational resources; they also demonstrate the ability to bend or flout the law. In this view, incumbents that manipulate elections extensively do not betray weakness in the form of unpopularity; instead they display alternative sources of resilience that can be used to compensate for low popular approval. Moreover, since these signals are costly, they can affect the behavior of other political actors [@gehlbach2015electoral]. I call this the regime-strength model. 

In this paper, I further develop the theory of a regime-strength model of electoral manipulation, and test it using new empirical approaches.  In particular, I argue that extensive electoral manipulation helps induce elite solidarity with the incumbent, if and when post-election protests occur, by two mechanisms. First (and most importantly), recruiting low-level actors to perform illicit electoral manipulation entails significant principle-agent problems. Extensive illicit electoral manipulation is, as a result, a costly signal of incumbents' ability to overcome those principal-agent problems through deep patronage resources and effective control over the security apparatus. Second, this costly indicator of 'strength'---deep pockets and influence over the police, prosecutors, and courts---acts as a coordination mechanism for elites during protest events. During a post-election challenge, the behavior of an individual elite (whether or not to remain loyal to the incumbent) will in part be conditioned on their expectation about *other* elite actors' behavior. The costly signal of resources and capacities produced by extensive illicit manipulation makes it appear more likely that other elites will remain loyal; a potentially self-fulfilling prophecy [@hale_patronal_2014]. Perversely, and counter to the expectation of much existing theory, this means incumbents have incentives to manipulate *more* heavily when they face greater opposition threat.

The stakes of this debate are significant for theories of democratization and autocratization. The regime-weakness view makes auspicious predictions for democratization and ultimately under-pins the notion of 'self-enforcing' democracy [@fearon_self-enforcing_2011]. If the latter view holds, the risk of electoral protest may fail to deter election manipulation; greater protest risk may instead incentivize incumbents to manipulate elections more profoundly in order to demonstrate their capacities to other actors in the system. In addition, a sizable literature has been founded on the regime-weakness framework. Scholars working in this mode have argued that manipulation can be kept in check, and ultimately reduced, through the development of institutions that reveal information about election integrity: election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting]. 

At first glance, it would seem straightforward to test these competing models by comparing estimates of electoral manipulation with the incidence of post-election protest. The regime-weakness model typically predicts that increasing manipulation (at a constant level of true popularity) increases the likelihood that opposition groups will initiate post-election protest [@Luo2017;@Little2012;@Schedler2002;@Beaulieu2014a]. By contrast, it might be presumed that the regime-strength model would yield the opposite prediction [@simpser2013]; that protest initiation decreases with the level of manipulation (holding vote-share constant). Yet empirical findings on the relationship between manipulation and protest have been mixed [@Hyde2014;@daxecker2012;@Hafner-Burton2016;@Lankina2017;@ananyevDictatorsSignalStrength2022]. 

There are many possible reasons for these mixed results. One fundamental problem with protest initiation as the dependent variable in this context, however, is the complex and multi-purpose nature of post-election protest. The predictions of the regime-weakness model hinge on the assumption that protest initiation is narrowly instrumental---driven by the expected benefit of overthrowing the incumbent. However, protest initiation may be driven by various actors with different motivations, including citizen grievances and the long-term goals of protest entrepreneurs (e.g. generating publicity, waging factional battles, costly signaling in bargaining for policy change, etc.). Some of these reasons hold even when the likelihood of ousting the incumbent appears remote, meaning that protest can co-occur with signals of incumbent strength. Others of these might confound the apparent relationship between manipulation and protest initiation. A highly unpopular incumbent might need extensive manipulation simply to win the election, but deep unpopularity is also likely to provide fertile ground for protests. In such a scenario, the apparent correlation between high manipulation and a high likelihood of protest could be confounded by citizens' grievances. 

This paper does not aim to untangle the thorny relationships among citizen grievance, protest entrepreneurs' motivations, electoral manipulation, and protest initiation. Rather, it seeks to sidestep them by employing novel tests of the regime-strength and weakness models of manipulation. I develop two modeling strategies. First, I estimate propensity scores for post-election protest in a cross-national sample of electoral authoritarian regimes using structural factors that precede the actual level of manipulation in the election. These propensity scores are then used to test how electoral manipulation responds to higher structural protest risk. Second, among the subset of election-years in which protest does occur, I model electoral protest *duration* as a dependent variable. Once protests run their course, the 'strength' of an incumbent (its likelihood of outlasting a protest movement) is no longer latent. Instead, the incumbent's response\textemdash repression, concessions, or attrition\textemdash and its survival (or not) reveal the incumbents' true strength relative to the opposition. The more quickly a government puts down protests, the greater the underlying (unobserved) distribution of power favors the incumbent over the opposition. By contrast, the longer it takes the government to restore control, the less favorable the distribution of power for the incumbent.

[UPDATE AFTER REVISIONS ->]
Using data from 659 non-democratic elections across 100 countries, from 1989 to 2012 (of which 211 experienced protest), this paper finds that governments that manipulate elections more heavily are likely to put down protests more quickly, especially when incumbents claim small vote-shares. This is inconsistent with the regime-weakness model of election manipulation, but consistent with a world in which election manipulation indicates sources of regime resilience. Given these results, it is likely that---while incumbents do not seek to provoke protest---the downside risk for incumbents that engage in election fraud in non-democracies has been overemphasized. This interpretation helps explain several empirical patterns at once: the relative paucity of electoral protests [@Brancati2016], the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory in conjunction with high levels of fraud. At a theoretical level, they also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations which emphasize the role of front-line actors in delivering manipulation. Finally, they suggest a pessimistic outlook for democratization by election in authoritarian regimes.

[INCORPORATE THIS FROM BOOK ->]
This interpretation offers a unified explanation for several secular patterns: the infrequency of electoral protests [@Brancati2016], the limited success of such protests in toppling incumbent regimes, and the tendency of governments to report implausibly large victory margins in heavily rigged elections. For theories of electoral manipulation, it implies that more attention should be given to alternative deterrents to manipulation (such as the role of risk to low-level actors who implement manipulation). Ultimately, this perspective points to a more pessimistic view of democratization through elections in authoritarian contexts, and for the logic of 'self-enforcing democracy.'



# Literature review: regime strength, weakness, and election manipulation


The literature on authoritarian elections, electoral manipulation, and protest has developed competing (and at times contradictory) assumptions about the nature of regime 'strength' [@leventogluSynthesizingTheoriesAuthoritarian2025]. Fundamentally, a strong regime is one with a high probability of surviving a non-electoral challenge like a mass protest event or an attempted coup [@leventogluSynthesizingTheoriesAuthoritarian2025]. Sources of strength may include the loyalty of the military and security services, patronage resources, the financial or security backing of foreign powers, and more. By surviving---and deterring---challenges, a strong regime is also likely to be *durable*, though even a strong regime can fail. 

More contentious is the relationship between the incumbent's popularity and regime strength. The incumbent's popularity is his likelihood of winning a free and fair election. Some influential models conflate strength with popularity; in these models, popularity determines whether an incumbent is likely to survive a challenge. In those models, electoral manipulation can only signal weakness if exposed, since manufactured votes displace authentic ones for any particular vote total [@leventogluSynthesizingTheoriesAuthoritarian2025]. Popularity is, in my view, another component of strength against extra-electoral challenges. A more popular regime can rely on its genuine supporters to counter-mobilize against protesters or coup plotters [@hellmeierPullingStringsStrategic2020], and popular goodwill can deprive its opponents of recruits [@rosenfeldAutocraticMiddleClass2021]. Having identified these conceptual distinctions, we now turn to the benefits and risks of election manipulation for incumbents.

## Authoritarian principals and elections

As discussed above, authoritarian leaders face two central threats: mass revolt from below and elite coups [@svolik_politics_2012]. Elections help incumbents manage this dilemma by providing information about citizens, opposition groups, and the popularity and effectiveness of subordinates [@Geddes1989;@Miller2015a;@magaloni2006;@blaydes2011], while also helping recruit ambitious politicians and manage intra-elite disputes [@lust_okar2006].

Of course, holding an election creates the possibility of *losing* an election. Early research emphasized manipulation as a tool for avoiding electoral defeat. From this perspective, incumbents have more to gain from election rigging when they are likely to narrowly lose the election without it [@nybladeWhoCheatsWho2008;@lehoucq2002;@kamSecretBallotMarket2017;@Birch2007;@Fukumoto2011], and when the prospect of losing carries significant costs [@ziblatt2009;@baturoDemocracyDictatorshipTerm2014;@acemoglu2005economic;@Schedler2013;@horzElectoralManipulationPolarized2021]. Subsequent research highlighted broader benefits. Excessive or blatant manipulation can signal regime resources and durability, encouraging others to cooperate with the incumbent [@simpser2013;@gehlbach2015electoral]. Costlier forms of manipulation may provide especially credible signals of strength and thereby reduce protest [@Harvey2020].

On the other hand, manipulation nevertheless entails substantial material and organizational costs. Even relatively cost-effective tactics like co-opting employers to pressure voters requires monitoring efforts and rewards for compliant businesses [@frye2014, p. 207]. Agents and brokers engaged in falsification or other types of manipulation must still be organized and compensated [@Langston2009mexico]. Vote-buying can be particularly costly [@lehoucq2002;@wang2007], with costs rising as elections become more competitive [@corstange2018clientelism]. Wang and Kurzman [-@wang2007], for example, estimate that vote-buying in a single Taiwanese county in 1993 could cost more than $4 million (1993 dollars), roughly twelve times the legal campaign expenditure limit. More broadly, manipulation draws on finite resources---including cash, goods, jobs, regulatory pressure, and favors---that must be traded off against other electoral or political purposes.

Beyond the organizational cost, scholars have also focused on the 'legitimacy cost' of election manipulation [@birch2011]. Legitimacy is understood as an individual's subjective sense that authorities have the right to rule, even if the individual may object to  particular decisions [@Lipset1960]. Greater legitimacy, in turn, makes it more likely that individuals will comply with authorities' demands [@Levi_Sacks2009]. Legitimacy cost, then, is both psychological and potentially behavioral [@Schedler2013]. Post-election protest is the most visible, and possibly most consequential, legitimacy cost of electoral manipulation [@birch2011]. 

This understanding of electoral manipulation and protest has been widespread in the literature. In several formal models of manipulation, loss of support and/or protest risk often stand as the main downside of manipulation for incumbents [@Little2015;@littleFraudMonitoringNoncompetitive2015;@chernykh2015third]. Protest risk as the cost of manipulation also figures in less formal theory. It is central to the literature on election monitoring [@hyde_2012_marinov;@kelley2012] Manipulated elections can serve as a focal point for citizens' diverse grievances, raising the likelihood of protest [@tucker2007;@higashijimaDictatorsDilemmaBallot2022].

In sum, when holding an election, the existing literature expects incumbents to try to optimize across several goals: to win the election, to send signals of 'strength' to other political actors, to manage organizational costs, and to minimize legitimacy costs (typically defined as the risk of protest). As a result, a theory of electoral manipulation must also account for the motivations of post-election protest.  Integrating agent risk more fully in the theory of manipulation will show that the most fundamental 'legitimacy cost'---the risk of losing office in a revolt---actually declines as electoral manipulation increases. 

## Post-election protest

Regime-weakness models assume that opposition protest is *elite-led and instrumental*. The opposition is typically modeled as a unitary, strategic actor who stages protest when the expected benefits of successfully overthrowing the incumbent outweighs the expected cost of failure. The signal of weakness sent by election manipulation, in turn, indicates to opposition actors that protest is more likely to succeed in unseating the incumbent [@Luo2017;@Little2012;@Schedler2002;@Beaulieu2014a]. Incumbents in these models typically face a tradeoff between the short-term benefits of electoral manipulation for winning the election, and any weakness that relying on fraud reveals to the opposition [@Rozenas2016b]. As Beaulieu [-@Beaulieu2014a, p. 48] writes in this tradition, "As the opposition grows stronger, the amount of [manipulation] he will tolerate shrinks, and the incumbent who wishes to avoid protest will need to reduce the amount of manipulation she commits." Such abstraction helps keep formal models tractable, but this assumption is difficult to square with recent research on electoral protest and a broader understanding of political protest in authoritarian regimes.

Electoral protest may occur because manipulation galvanizes citizens who hold diverse grievances [@tucker2007], because economic conditions are poor [@Brancati2016;@Beaulieu2014a], because electoral manipulation makes people angry [@Szakonyi2021], or simply because their party lost the election [@andersonLearningLoseElection2006]. Opposition elites, in turn, might seek to initiate---or amplify---protest to take advantage of this popular discontent for broad, rather than instrumental, reasons. They may wish to pressure the incumbent in order to push for improved competitiveness in the *next* election or to bargain for policy concessions [@chernykh2014;@robertson_protest_2010]. They may stage protests in order to identify new participants, forge new social ties [@barberaModelProtestsRevolution2019] and test new tactics. They may seek to communicate information about the government's unpopularity, to set the stage for larger, future protests [@lohmannDynamicsInformationalCascades1994]. Protests may be staged as part of internal factional rivalries within the opposition [@Buckles2017], or to drive media coverage of issues important to the opposition [@huangMediaProtestDiffusion2019;@jenningsStreetDemonstrationsMedia2019]. Indeed, opposition elites may take advantage of the bare fact of an election to challenge the incumbent, irrespective of the level of manipulation or the incumbent's vote-share [@Pop-Eleches2015]. If protest is dispersed and multipurpose, protest movements may initiate for reasons orthogonal to perceived incumbent strength and severity of electoral manipulation. In such a setting, incumbents cannot count on quiet streets simply by manipulating at a low level, or even holding clean elections.

The varied motivations for protest may help account for the inconsistent relationship between election manipulation and protest initiation found in the literature. Several cross-national studies have found links between election manipulation and binary measures of protest [@Brancati2016;@Harvey2020;@Hyde2014]. A study of African elections from 1997 to 2009 shows that election fraud is associated with an increased number of post-election conflicts [@daxecker2012]. Similar results are found by Rød [-@Rod2019], using a different measure of manipulation.  Using an alternative measure of manipulation from the NELDA dataset, however, Hafner-Burton et al [-@Hafner-Burton2016] find that the association between election fraud and post-election protest falls short of conventional statistical significance. Studies of one archetypal example, the 2011 election and protest cycle in Russia, have found a positive relationship between fraud and protest by some measures [@Lankina2017], and no relationship by others [@ananyevDictatorsSignalStrength2022]. As noted earlier, Simpser finds a negative relationship between the severity of election manipulation and the likelihood of protest [@simpser2013] and a positive relationship between 'excessive manipulation' (i.e. extensive manipulation and a large incumbent vote-share) and regime tenure. 

[ADD NEW SEGUE TO THEORY SECTION HERE]



<!-- NOT SURE WHERE THIS GOES:
Empirical work has found some support for the microfoundations of this theory. For example, ballot-stuffing and voter coercion have been shown to generate stronger feelings of anger than pre-election manipulation of ballot access [@Szakonyi2021], and access to information about manipulation appears to reduce support for the ruling party [@Reuter2021], increase support for protest and reduce the perceived legitimacy of the regime [@Williamson2021]. This argument has found support in studies of the relationship between economic grievances, manipulated elections, and protest [@Rod2019;@Brancati2016].
-->

# Theory: electoral manipulation, incumbent strength, and protest duration

BOOK TEXT:

There are two major dimensions to this argument. First, I argue that the same capacities that incumbents rely on to manipulate elections---patronage resources and impunity for agents---are useful for defeating protest movements. Consequently, incumbent regimes that manipulate elections extensively are intrinsically likely to have the tools needed to survive post-election protests. This means that incumbents who are capable of extensive manipulation face reduced legitimacy costs from that same manipulation; they are more likely to survive the possible backlash. 

Second, expanding upon Simpser's work [-@simpser2013] I argue that extensive electoral manipulation sends signals about these resources and capacities to elites, which incentivize them to cooperate with the incumbent during protest movements. Because it requires the resolution of significant principal-agent problems, extensive electoral manipulation is a costly signal of the incumbent's patronage resources, control over the security environment and the courts, and ability to command agent loyalty broadly across the territory of the country. It further signals the regime's intention to devote resources within the elite selectorate, rather than more broadly among society. The more extensive the manipulation, the stronger these signals in favor of elite solidarity are likely to be. 

This model extends Simpser's work in two ways. First, it grounds the costly nature of the strength signal in the principal-agent dynamic; if manipulation could simply be willed into existence by the incumbent, it would not be costly [@Harvey2020]. By overcoming a huge set of principal-agent problems, the incumbent 'pays' to display his patronage capacity and influence over agents. This is a fundamental reason why electoral manipulation is not cheap talk, as some deterrence models posit. Second, I argue that these signals do not require manipulation to be extensive or blatant, as in Simpser [-@simpser2013]. To be successful in this regard, manipulation need not result in a wide margin of victory or be carried out in plain view; it is only necessary that the manipulation effort be intelligible to elites themselves.

## Electoral manipulation, principal-agent problems, and costly signaling


Extensive electoral manipulation requires cooperation from a large network of subordinates who bribe or pressure voters, stuff ballot boxes, or falsify results [@rundlett2016deliver]. These agents may include partisan activists [@stokesBrokersVotersClientelism2013], election officials [@baderRolePrecinctCommissions2017;@Sjoberg2016], neighborhood brokers [@langstonWhyClientelistBrokers2023], local mayors and civil servants [@maresCoreVotersCurse2018], public employees [@Forrat2018], and workplace managers [@fryeWorkplacePoliticsHow2025;@cohnpostarVoteYourBread2021].

Mobilizing thousands of such individuals poses substantial principal-agent problems. Whereas early research often assumed that incumbents' orders were smoothly implemented, subsequent research has emphasized the risks posed by subordinates [@svolik_politics_2012;@svolikPowerSharingLeadership2009;@magaloni2008credible], pervasive principal-agent problems [@rundlett2016deliver;@Harvey2019;@tysonAgencyProblemUnderlying2018], and difficulties monitoring low-level agents [@Larreguy2016;@Boix2013]. These networks are typically pyramidal, with each level overseeing a larger number of actors below it [@hale_patronal_2014]. Yet agents may not behave as principals intend [@rundlett2016deliver], and actors throughout the hierarchy have incentives to misrepresent their loyalty and effort [@harveyWhyAllowLocal2022]. Large-scale manipulation therefore depends on a hierarchical network in which high-level actors such as party leaders, governors, and business owners influence meso-level actors such as mayors and firm directors, who in turn direct large numbers of front-line agents such as election officials, civil servants, and party operatives.

Agents' incentives also diverge fundamentally from those of political principals. Principals benefit directly from manipulation because it increases their chances of winning close elections [@lehoucq2002] and can widen their margins of victory [@greene_why_2007;@magaloni2006;@gehlbach2015electoral;@simpser2013]. Agents, by contrast, are primarily motivated by access to patronage---jobs, rents, and other benefits associated with political networks [@diaz2006federalism;@lust_okar2006;@reuter2012subnational]---and by the risk of exposure and punishment for breaking the law [@Harvey2022].


Consequently, incumbents who produce extensive electoral manipulation must have deep patronage resources [@Harvey2019;@hale_patronal_2014;@greene_why_2007] and the ability to shield agents from the consequences of exposure. By overcoming a large number of nested, geographically dispersed principal-agent problems, an incumbent has produced a costly signal of those resources and capacities. Crucially, many of the same resources and capacities improve the incumbent's ability to respond to a nascent protest movement through repression or concessions [@tillyMobilizationRevolution1978;@lipskyProtest1968;@Hummel2019;@raslerConcessionsRepression1996;@Bishara2015;@Frantz2014]. Patronage resources may be aimed at protesting parties themselves [@reuter2015;@mooreRepressionDissentSubstitution2000;@Hafner-Burton2016;@Piven1979;@robertson_protest_2010] or the more moderate public, in order to keep them on the sidelines of the conflict [@Hummel2019]. Incumbents may also threaten to withhold resources from clients in exchange for passive support [@Rosenfeld2017], sapping the pool of potential protesters.   

Ruling parties may also lean on repression, calling upon the security services to disperse protests, arrest leaders, and perhaps employ harsher methods including disappearance, torture, and execution [@Escriba_Folch2013;@Davenport2007;@DeVogel2022]. The incumbent's ability to devote patronage resources to the security services helps maintain their loyalty during such moments, by suggesting that the security forces have more to lose if the incumbent is ousted [@bellinRobustnessAuthoritarianismMiddle2004;@Albrecht_Ohl_2016]. The ability to protect regime agents from any legal consequences for carrying out orders, a condition which helps enable illegal election manipulation [@Harvey2022], is also an asset when asking law enforcement agents to carry out possibly violent repression [@MAGALONI_RODRIGUEZ_2020;@stavro_police_2023;@abouharbFactoJudicialIndependence2013]. 

This is "infrastructural power" [@leventogluSynthesizingTheoriesAuthoritarian2025]---those sources of regime durability that derive from its ability to mobilize and discipline agents, to distribute patronage, and to deploy coercion. This contrasts with the legitimacy-based view of strength centered in the regime-weakness framework. As a result, I expect that large-scale electoral manipulation will positively influence elites' expectations of the regime's long-term durability, helping prevent elite splits during a post-election challenge [@BuenodeMesquita2005;@svolik_politics_2012].




As articulated by Simpser [-@simpser2013], electoral manipulation can communicate information about the resources and capacities of the incumbent ruling party. This information can then persuade other political actors to remain loyal to the incumbent, rather than join a challenge. Incorporating the principal-agent dynamics discussed in this book shows why electoral manipulation need not be excessive or blatant to serve this function. In Simpser's model, other political actors learn about the level of electoral manipulation through its public display and/or by observing an implausibly high margin of victory for the ruling party. The model articulated here shows how manipulation efforts require the cooperation of large numbers of people, including elites within the regime.^[Regional governors, big business owners, and directors of security forces are all examples of elites whose resources can be mobilized at election time.] Internal elites observe their own level of manipulation effort, and can estimate those of their peers [@rundlett2016deliver]. An extensive electoral manipulation effort can therefore be covert to the public, while intelligible to elites. Knowing the difficulty of mobilizing agents to engage in illicit manipulation, such an effort signals regime strength and should prevent splits from within the elite.

## Regime strength and protest duration

Considering how electoral manipulation may be tied to protest duration under the regime-weakness and regime-strength models reveals distinct predictions. Specifically, when manipulation increases, protest movements should last longer under the regime-weakness model, but be defeated more quickly under the regime-strength model (conditional on incumbent vote-share).^[In both cases, this assume that incumbent survives the protest movement, by far the modal outcome. When the opposition wins, an outcome considered in the appendix, the assumption would be reversed: a quicker opposition victory indicates a weaker incumbent.] The logic hinges on the assumption that, on average, authoritarian governments will seek to quash protest movements quickly once they appear. How plausible is this assumption?


Mass protest is one of the two primary threats to authoritarian regimes [@svolik_politics_2012], and as such they often go to great lengths to prevent it. These efforts include costly (and risky) investments in the repressive apparatus [@Escriba_Folch2013;@draguPreventiveRepressionTwo2019], co-optation of opposition parties [@reuter2015], and large-scale monitoring and censorship schemes [@King2013]. And they often relentlessly hound would-be protest entrepreneurs to deter such behavior [@xuRepressCooptAuthoritarian2021a;@mehrlSecretPoliceOrganizations2024]. Mass protest then, is clearly undesirable from the authoritarian incumbent's point of view, and we can plausibly assume that once an electoral protest movement begins, the incumbent will exert considerable effort to quash it. Governments that are stronger, relative to the protesting opposition, will defeat protest movements faster than will weaker governments. As argued below, electoral manipulation can be demonstrative of capacities that may also be used to defeat protest movements.

An important objection could be raised here. Might not a stronger incumbent, feeling secure in their power, be willing to tolerate opposition protest longer than a weak incumbent? It is true that incumbents do sometimes tolerate protest movements [@yuenNeitherRepressionConcession2017]. However, there are two reasons this is unlikely to be the case systematically here. First, authoritarian governments are more likely to tolerate protests aimed at low-salience, local issues [@lorentzenRegularizingRiotingPermitting2013;@kleinDynamicsPoliticalProtests2018], and more likely to clamp down on protests that have higher concession costs [@kleinDynamicsPoliticalProtests2018]. As electoral protests challenge the incumbent's ostensible victory and right to remain in office, they are directly threatening to the regime; the concession costs of stepping down or holding new elections are very high. Second, authoritarian leaders are working in an information-poor environment [@guriev2020], in which mass protests can unexpectedly grow from small sparks to wildfires that quickly sweep away the regime [@Kuran1991]. "Ignoring" protests can lead to further mobilization against authoritarian regimes [@Bishara2015], as can concessions in the absence of durable political reform [@leuschnerStateConcessionsProtest2024]. In such cases, we should expect to see apparent 'toleration' displaced by 'attrition'---efforts by the incumbent to increase the cost of protest while avoiding open repression [@yuenNeitherRepressionConcession2017]. As a control, however, I include a measure of average protest size in one model of protest duration, in the expectation that larger protests are harder to tolerate. 


Under what circumstances should protest movements be expected to last longer in authoritarian regimes? Authoritarian incumbents have strong incentives to keep protests small and short-lived. Generally speaking, by definition, authoritarian regimes have a smaller winning coalition and selectorate than democracies do, and allocate more private and club goods to elites as a result [@BuenodeMesquita2005]. Despite this, they can appear superficially popular, with most citizens falsifying their preferences and demonstrating support for the regime, tacitly or otherwise. The appearance of even a moderate number of protesters publicly signalling their disapproval of the government can trigger a cascade of dissent in short order, as other citizens stop falsifying their preferences as well [@Kuran1991]. Given time, persistent protests can lead elites and the security services to update their prior beliefs about the strength of the incumbent, prompting defections [@svolik_politics_2012;@Chenoweth_Stephan2011]. Protest movements can also challenge the regime's legitimacy [@keremouglu2022thin], eroding individual compliance in other areas. To prevent all these negative outcomes, incumbents try to prevent mass protest from occurring, and to quickly quell those protest movements that do arise.

Assuming the incumbent ultimately remains in office, stronger protest movements will tend to last longer than weaker ones. Ordinary protesters are more likely to remain engaged when they are more aggrieved [@gurrWhyMenRebel2015], when they are angrier, and when they observe many other people expressing their discontent; extensive public grievance can thus be a source of durability for protest movements. The longer-lasting and larger protest movements are, the more protesters may become socially networked, enabling easier subsequent mobilization [@schussmanProcessProtestAccounting2005;@sheldonIdentityCompetitionKey2000], increasing organizational resilience and capacity . When group goals appear more achievable---in part because the protest movement is large, well-organized, and well-resourced, and the incumbent appears off balance---individuals we will be more willing to participate in protest [@Klandermans1997] and to become regulars [@saundersExplainingDifferentialProtest2012].  Protest entrepreneurs, who are using protest as a form of costly bargaining against the incumbent, will aim to demonstrate the resilience and popularity of their movements in order to present a stronger negotiating position [@robertson_protest_2010;@Beaulieu2014a]. As a result, they often seek to maintain high levels of mobilization even in the face of repression and ordinary citizens' waning interest over time [@tillyContentiousPolitics2015]. On average, then, I expect that authoritarian and proto-authoritarian incumbents will seek to demobilize protest, and protest movement leaders will seek to maintain (or even increase) high levels of mobilization. Which side prevails, and how quickly, depends on the relative strength of each.

Electoral manipulation has been shown to be more extensive under conditions that minimize that minimize principal-agent problems (i.e. those that make low-level agents more likely to cooperate with the incumbent). These include: greater local popularity for the incumbent [@rundlett2016deliver], greater incumbent control over patronage resources [@Harvey2019;@hale_patronal_2014], and lower risk of punishment for pro-regime lawbreaking [@Harvey2022]. As discussed below, these capacities and resources are also independently useful for surviving post-election protest. Consequently, if an incumbent is able to manufacture a large numbers of votes, it is also likely to have the resource capacity, control over bureaucratic institutions, and perhaps deep wells of popularity in parts of the country that can serve as independent sources of regime durability, even if the incumbent's nationwide margin of victory is low. Lastly, as I have argued in this volume, by overcoming principal-agent problems arising from agent risk, extensive illicit manipulation also indicates that large number of agents assess that the incumbent will remain in power in the future [@simpser2013]. This can itself become a self-reinforcing equilibrium [@hale_patronal_2014].

Still, a full accounting of a regime's strength would include the incumbent's "infrastructural power" *and* their genuine popularity. Like infrastructural power, real popularity among some constituencies should help the government survive a post-election challenge [@Gerschewski2013]. Of course, in most authoritarian cases, the incumbent's true popularity is unknown due to censorship, repression of opposition candidates, and preference falsification by citizens. As a result, scholars (and political actors in real time) are left to use the incumbent's vote-share and perceived level of manipulation as imperfect indicators of the incumbent's true popularity. For example, an election in which the incumbent records a high vote-share with limited electoral manipulation might be relatively popular while lacking the capacity to manipulate elections extensively.^[Such a case might also indicate an incumbent with high capacity to manipulate who *chooses* not to do so, a prospect considered in the next section.] If a post-election challenge were to occur here, the incumbent's low infrastructural power (e.g. limited patronage resources and weak control of the security forces) might be offset by its relatively high popularity, enabling the government to survive the challenge. Likewise, a low incumbent vote-share in a highly rigged election indicates low popularity but high infrastructural power; the risk of losing a challenge due to low popularity may be offset by the incumbent's other resources and capacities. Furthermore, even genuinely popular incumbents may resort to large-scale manipulation in order to bolster turnout [@Smyth2018;@martinezicomaElectionTurnoutAuthoritarian2020]. Since genuine popularity cannot be directly measured, but is indicated by the mix of incumbent vote-share and level of manipulation, both indicators must be included in our hypotheses. 


Putting it all together: this theory assumes that authoritarian incumbents generally prefer order to unrest, and that protest leaders want to demonstrate the popularity of their cause through large, sustained protests. Electoral protests are a form of competition between the incumbent and opposition actors, in which both sides are trying to demonstrate their strength in order to win concessions from\textemdash or defeat outright\textemdash the other [@Beaulieu2014a]. The outcome of this contest reveals the underlying strengths of the participants. Consequently, governments that quickly defeat protest movements can be considered stronger than those that take weeks or months to prevail, which in turn can be considered stronger than those governments which collapse in the face of mass protest.  Incumbents who manipulate extensively demonstrate they have resources and capacities that can also be used to buy off, repress, or otherwise quash post-election protest; they also send costly signals of such to elites that can induce their loyalty. Such signals of extra-electoral strength should be especially informative to elites when regime popularity is low. This yields Hypothesis 1.

> H1: Election manipulation is negatively associated with protest movement survival, with the size of this effect decreasing with incumbent vote-share.

[ALTERNATIVE VERSION:]

>**H1 (Unpopularity offset by instractural power)**: Increases in electoral manipulation will be associated with decreasing duration of post-election protest when incumbent vote-share is low.



All told, the core of the argument is as follows: incumbents generally aim to manipulate elections as extensively as possible to reap both the direct and indirect benefits of doing so [@simpser2013]. Achieving this requires overcoming principal-agent problems---by tightening control over patronage networks and insulating agents from potential consequences. The very resources that enable large-scale manipulation—such as genuine popularity, control over state resources, and dominance of the legal system---also equip incumbents to suppress or dismantle protest movements when they arise. Consequently, more extensive manipulation should correlate with shorter durations of anti-incumbent protests.


This argument\textemdash that successful, large-scale election manipulation draws on the same resources a state may use to suppress protest\textemdash turns the deterrent logic on its head. Instead of signalling weakness, large-scale fraud indicates a well of resilience: control over resources, impunity for pro-regime lawbreaking, and loyal agents. As a result, governments that build and deploy an electoral manipulation apparatus are simultaneously inoculating themselves against the expected cost of protest.

[OLD TEXT:]

The central argument of this manuscript is that, in authoritarian regimes, incumbents who develop the tools necessary to manipulate elections extensively will also have the tools necessary to more quickly prevail over protest movements. This proposition builds off Simpser's foundational argument that manipulated elections can send costly signals of the incumbent's "capacities, resources, [and] dispositions" that will help it retain power in the future [@simpser2013, p. 80]---that is, of its political strength. However, I depart from Simpser in two ways. First, by incorporating more recent research on the principal-agent dynamics inherent in election manipulation, I argue that incumbents who can produce large amounts of election manipulation necessarily have "capacities, resources and dispositions" helpful for defeating protests---a peripheral subject for Simpser. Second, mine is not a signalling argument with regard to protesters (though the election outcome may still communicate signals of regime strength to, for instance, state agents deciding how to respond to the protest movement). I argue that electoral protest may occur for many reasons, even when incumbents are perceived to be 'strong.' However, good manipulators are likely to have deeper wells of strength than ineffective manipulators do, even when they fail to 'deter' post-election protest. This is due, in part, to the regime resources and cohesiveness that widespread election manipulation requires, and in part due to the fact that widespread manipulation is a costly signal of such capacities. 


It is likely that the theoretical mechanisms described above weaken as regimes become more liberal-democratic, as Simpser notes [-@simpser2013, p. 192]; in liberal democracies, patronage consolidation should decline and rule of law should increase by definition. These changes should make it more difficult for incumbents to manipulate elections, irrespective of protest risk. At the same time, the more liberal-democratic a state becomes, the more alternatives there are to protest as a means to check the executive. More independent courts and more active executive constraints have been shown to be associated with improved election integrity [@Harvey2022;@Bishop2016a]; it is where these factors are weakest---the electoral authoritarian regimes---where we should expect the deterrence logic to operate under a regime-weakness model. Finally, as a matter of measurement, protest duration is likely to be an increasingly imperfect measure of incumbent strength in more liberal-democratic settings, where protest is more commonplace and less immediately threatening to the regime.

This model claims that more extensive illicit manipulation serves as a costly signal or regime strength that helps induce elite loyalty during post-election challenges. As such, it also makes a distinctive prediction about how incumbents should respond to the possibility of post-election contestation. Rather than back down in the face of protest risk, as regime-weakness models predict, the regime-strength model suggests that incumbents should work to intensify manipulation. By doing so, they can send a more convincing signal of their staying power to elites, preventing defections, and ultimately *reducing* the risk that the incumbent will lose office.

>*H2 (Signaling to induce loyalty)*: Electoral manipulation should increase with pre-election indicators of protest risk.


In other words, the theory predicts that *incubments should manipulate more when they face strong oppositions*, rather than moderate their behavior, as conventional wisdom might hold, and that when post-election protests do occur, *greater electoral manipulation will be associated with quicker defeat of protest movements.*

# Data, case selection, and main variables


These hypotheses are tested using data from three datasets: NELDA [@hyde_2012_marinov], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019]. The unit of observation for this study is an election period as recorded in NELDA. This prevents selecting on the dependent variable, by capturing both protested and non-protested elections. I limit the sample to authoritarian regimes which hold multiparty national elections as coded in V-Dem. This excludes fully closed autocracies (where no national elections are held), as well as electoral and liberal democracies. I limit the sample to those regime-years coded 1 (electoral authoritarian) on the V-Dem variable 'v2x_regime' for the election year, or coded 2 (electoral democracy) for the election year but 1 for the lagged year.^[This coding rule will capture cases that are coded as electoral democratic after an election but electoral authoritarian before, in order to ensure that cases which experience a liberalization after the election are included.] This sample includes a range of electoral authoritarian cases, including those that might be labeled as 'hegemonic' and 'competitive' [@donno2013elections].^[According to the V-Dem codebook, the coding scheme includes cases where multiparty elections were manipulated to the extent that "irregularities...affected the outcome of the election," and ones in which it is "hard to determine whether the irregularities affected the outcome or not." Only when irregularities definitively did not affect the outcome would election-years be categorized as electoral democracy, and excluded from the main analysis.] For instance, the dataset includes election years from Egypt and Belarus, but also from Ukraine, and Mexico in the 1990s.

The case selection strategy here, excluding democracies, is one that both fits the theory and makes for a stronger test of the regime-weakness model. Nevertheless, an expanded sample that includes a set of electoral democracies provides similar results in the appendix. In total, there are 659 elections included in the dataset, across 100 countries, with 211 experiencing post-election protest. The date range is limited by the availability of data on protest from ECAV, which ranges from 1989 to 2012.

To capture the severity of *election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." It is important to note that this measure captures illegal election-day manipulation broadly, and is not confined simply to falsification. 

I operationalize electoral manipulation using this variable, rather than alternative measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], may not mobilize real voters [@Harvey2020], and can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021]. This measure of election integrity should thus serve as a strong test of the regime-weakness model: if manipulation indicates incumbent weakness, it will be most likely to do so when employed in this way. It also matches the conception of election manipulation used in much of the relevant formal literature.^[For example, election manipulation is treated as post-hoc adjustment of the results in Little et al [-@Little2015], Little [-@Little2012], Magaloni [-@magaloni2010], and Fearon [-@fearon_self-enforcing_2011], among others.] 

Data on protest occurrence and duration are derived from ECAV, which records protests and other events related to an election, ranging from six months before to three months after the vote (or three months after the final round of a multi-round election). To qualify, an action must be contentious; rallies in support of the incumbent are excluded. I only include events labeled as protests, occupations, or blockades, excluding acts like shootings, bombings, and coups.


# Analysis 1: Protest risk and severity of manipulation

```{r setup, include = FALSE}
library(tidyverse)
library(colorspace)
library(countrycode)
library(readxl)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
library(stargazer)
library(ggeffects)
#library(glmmTMB)
library(modelsummary)
library(huxtable)
library(simPH)
library(broom)
library(broom.helpers)
library(kableExtra)
library(flextable)
library(gt)

nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2024-02-06.csv"))


nelda2.sub <- nelda2.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
nelda2.sub <- nelda2.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))



##Naming nelda variables

nelda2.sub <- nelda2.sub %>% rename(boycott = nelda14)
nelda2.sub <- nelda2.sub %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
                                                                                 NA)))

##Creating economic crisis variable

nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = as.factor(econ.crisis))

nelda2.sub <- nelda2.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))

nelda2.sub <- nelda2.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))



## Removing cases where inc. loses
#nelda2.sub <- nelda2.sub %>% filter(inc.lose == 0)  #See how this affects things

##Intmon as a factor
nelda2.sub <- nelda2.sub %>% mutate(v2elintmon = as.factor(v2elintmon))


nelda2.sub <- nelda2.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))

#Square of physint

nelda2.sub <- nelda2.sub %>% mutate(physintlag.sq = physinteg.1lag^2)

## Categorical variable for incumbent status

nelda2.sub <- nelda2.sub %>% mutate(incumb.status = as.factor(ifelse(inc.termlimit == "yes" & chosen.successor == 1, "Chosen successor",
                                                           ifelse(inc.termlimit == "yes" & chosen.successor == 0, "Contested exec. election", 
                                                                  ifelse(inc.termlimit == "no", "Executive eligible", "Legislative election"))))) 
nelda2.sub <- nelda2.sub %>% mutate(incumb.status = relevel(incumb.status, ref = "Legislative election"))


#Coding as factors

nelda2.sub <- nelda2.sub %>% mutate(presidential = factor(presidential))
#nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = factor(inc.termlimit.num))
nelda2.sub <- nelda2.sub %>% mutate(rentierstate = factor(rentierstate))
nelda2.sub <- nelda2.sub %>% mutate(taxstate = factor(taxstate))
nelda2.sub <- nelda2.sub %>% mutate(unscheduled_election = factor(unscheduled_election))
nelda2.sub <- nelda2.sub %>% mutate(inc.lose = factor(inc.lose))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))


#Successful protest = new elections or new incumbent

nelda2.sub <- nelda2.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))



# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections

nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))

## Replacing NA with 0 for protest size

#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(is.na(max_size)==T & n.events.post == 0, 0, max_size))

## Fixing missing values of winner.share.upd for liberal democracies

nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(is.na(winner.share.upd)==T & v2x_regime == 3, winner.margin, winner.share.upd))

## Set max_size zero for elections with no protest

#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))

## Excluding liberal democracies

nelda2.sub <- nelda2.sub %>% filter(v2x_regime == 1 | (v2x_regime == 2 & regime.1lag == 1)) #Includes only authoritarian, add e-dems with | v2x_regime == 2  #Current version includes lagged 1 in order to capture pre-election auth.

## Remove outlier
nelda2.sub <- nelda2.sub %>% filter(d.events.num <= 100)

#Protest cases only

protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1)

## Coding median size as 0 when n.events = 0

nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(n.events.post == 0, 0, median.participants))

## Adding other dimensions of manipulation

nelda2.sub <- nelda2.sub %>% mutate(v2elintim.inv = v2elintim * -1)
nelda2.sub <- nelda2.sub %>% mutate(v2elvotbuy.inv = v2elvotbuy * -1)
nelda2.sub <- nelda2.sub %>% mutate(emb_control = v2elembaut * -1)


nelda2.sub <- nelda2.sub %>% mutate(illegal_man_index = v2elintim.inv + v2elvotbuy.inv + v2elirreg.inv)

## Inc lose
nelda2.sub <- nelda2.sub %>% mutate(inc.lose_clean = as.factor(ifelse(inc.lose == "<NA>", NA, inc.lose)))
```



## Modeling strategy

I utilize two complementary approaches to test the relationship between protest risk and manipulation. The first of these takes the *level* of election manipulation as the outcome variable, while the second uses *changes* in the level of manipulation. Both employ two-stage models. 

In the first approach (where level of manipulation is the ultimate outcome variable), a logit model is used in the first stage to generate a propensity score of post-election protest in each case; the first-stage model uses only pre-election variables as predictors. The goal of this step is to estimate the underlying likelihood of post-election protest in each country-year, independent of the level of election fraud observed on election day. The fitted values (i.e. predicted probabilities of protest) for each observation are obtained as propensity scores for protest, to be used as explanatory variables in the second-stage models. The second-stage models take the level of *electoral fraud* as dependent variables. Deterrence models would expect that a higher propensity for protest would be associated with lower levels of manipulation, while the 'regime-strength' approach predicts the opposite; incumbents should work to intensify manipulation when protest risk is higher in order to signal strength to elites. 


It is important to clarify that I am not implementing a propensity-score *matching* or weighting approach here. Actual post-election protest cannot be considered a cause of electoral manipulation which occurs *before* the protest, and thus cannot be operationalized as a treatment variable. Instead, I aim to estimate the latent protest risk in each case, and analyze that risk as a predictor of observed electoral manipulation. 

**Stage one predictors** The effectiveness of this propensity score approach hinges on the success of the first-stage model, which should include variables that influence the likelihood of treatment. The first-stage model here includes lagged *judicial independence* [@Linzer2015], which can influence parties' decision to challenge elections in the streets vs. the courtroom. It also includes several lagged control variables taken from V-Dem, including measures of civil society organization ('v2cseeorgs') and legislative constraints on the executive ('v2xlg_legcon'), as estimates of the degree of contestation in the political system. The lagged openness of the media environment ('v2x_freexp_altinf') is included, since this could influence citizens' ability to gather information about the quality of elections and to communicate their grievances. De facto protection of individual physical integrity is included (lagged) along with its square term ('v2x_clphy'), to account for the linear and curvilinear effects of repression on protest. Lagged GDP growth ('e_migdpgro') is included as a measure of economic grievances. Finally, data on urban population share (from the World Bank) is included, since population density is a common correlate of protest. 

Three binary variables capture relevant features of the election, taken from NELDA. *Presidential* elections may have higher stakes, and thus be more prone to sparking protest. *Off-schedule elections* may be strategically timed by incumbents to account for more elevated protest risk. In addition, whether the executive is facing *term limits* is also included, since this can lead to fragmentation in the elite that can in turn yield protest [@hale_patronal_2014,@robertson_protest_2010]. The occurrence of a successful regional electoral protest in the previous year is included, derived from the ECAV data, given the potential to inspire protests in neighboring countries. 

Lastly, two final---and very important---covariates are included. These are: the number of pre-election protests (a very direct measure of protest risk), and lagged overall election integrity (i.e. overall manipulation in the previous national election, 'v2elfrfair' from V-Dem). 

```{r echo=FALSE, message=FALSE, warning=FALSE}
### Creating dataframe with fitted values for fraud

m.fraud.pred_base <- lm(v2elirreg.inv ~ n.events.pre + regional_success_lag, data = nelda2.sub)

m.fraud.pred <- lm(v2elirreg.inv ~ lji.lag +  gdpgro.1lag + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag  + v2elembaut + country, data = nelda2.sub)

#summary(m.fraud.pred)

nelda2.sub.fraudpred <- nelda2.sub %>% mutate(.rownames = rownames(nelda2.sub)) #id numbers for observations

new_frame <- augment(m.fraud.pred)
new_frame <- new_frame %>% dplyr::select(.fitted, .rownames, .resid) #fitted as residuals for each row

nelda2.sub.fraudpred <- nelda2.sub.fraudpred %>% left_join(new_frame, by = ".rownames") #merge back into fraudpred dataset
```



```{r, include = F, echo=FALSE}
### Calculate propensity scores for protest, but use lagged electoral fairness as variable

complete_cases <- complete.cases(nelda2.sub.fraudpred[, c("any.ecav.protest", "lji.lag" , "legcon.1lag" , "gdpgro.1lag" , "altinf.1lag"  , "frefair.1lag"  , "osorg.1lag"   ,   "presidential" , "inc.termlimit.num"  ,  "physinteg.1lag" , "physintlag.sq" , "urban.pct" , "n.events.pre" , "unscheduled_election" , "regional_success_lag", ".fitted", "country")]) #all obs in fraudpred that are complete for these variables

model.protest.null <- glm(any.ecav.protest ~ 1, family = binomial(link = "logit"), data = nelda2.sub.fraudpred[complete_cases, ])
summary(model.protest.null)

model.protest.nofraud <- glm(any.ecav.protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  + frefair.1lag  + osorg.1lag   +       + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + n.events.pre  +  unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub.fraudpred[complete_cases, ])
summary(model.protest.nofraud)

nelda2.sub.fraudpred$propensity_score <- NA

# Assign scores only to complete cases
nelda2.sub.fraudpred$propensity_score[complete_cases] <- predict(model.protest.nofraud, type = "response")

propensity_scores <- predict(model.protest.nofraud, type = "response")

```




```{r stage1-model, echo=FALSE, message=FALSE, warning=FALSE}
models4 <- list()
models4[['Model 1']] <- model.protest.nofraud
models4[['Model 2']] <- m.fraud.pred



row <- tribble(~term, ~Fraud, ~Residual,
               "", "Protest (logit)", "Fraud (OLS)",
                      "Country FE", "Y", "N")
attr(row, 'position') <- c(0, 30)



modelsummary::modelsummary(models4, gof_omit = "BIC|Log", longtable=F, stars = T, title = 'Stage 1 models', coef_map = 
               c("(Intercept)" = "(Intercept)", "presidential1" = "Presidential election", "lji.lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", 
                 "inc.termlimit.num" = "Incumbent term-limited", 
                 "urban.pct" = "Urbanization", 
                  "unscheduled_election" = "Off-schedule election",
                 "frefair.1lag" = "Overall elec. integrity (lag)",
                 "n.events.pre" = "Number of pre-election protests", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion"
                 ), add_rows = row) 
```

The results of the first stage model (with protest initiation as the DV) are presented in Table \@ref(tab:stage1-model). This model appears to capture the latent risk of protest reasonably well; the average propensity score for country-years where no protest ultimately occurred is 0.12. By contrast, the average propensity score for country-years where protests *did* occur is 0.66. 

In the second approach, I first model the average level of manipulation for each country, using only structural variables unrelated to protest risk. The residuals from this model---the observed deviation from the average for each country-year---can be interpreted as a measure of over- or under-production of manipulation relative to each country's baseline. I regress these residuals on of pre-election predictors of protest. Where the first approach estimates the level of fraud, the second approach estimates change in levels over time. As before, a deterrence model would predict that higher protest risk is negatively associated with these residuals (i.e. with *under*-production of illicit manipulation compared to baseline), while the regime-strength approach predicts the opposite. 


**Stage 1 predictors** The stage one, OLS model uses only structural variables to predict baseline manipulation. Country fixed effects are included so that this baseline level of fraud is isolated by country. Factors associated with classic protest deterrence, like the number of pre-election protests, are excluded. This should produce an estimate of the structural level of manipulation the incumbent could produce in the absence of strategic deterrence. The predictors include a V-Dem measure of the autonomy of election management bodies ('v2elembaut'), since less autonomous EMBs could be associated with higher levels of manipulation. A similar logic applies to the professionalization of the national bureaucracy ('v2stcritrecadm') and the degree of state control over the economy ('v2clstown'). Since access to and the source of government resources affect manipulation capacity, I include categorical variables for states that are primarily financed through taxation or resource rents ('v2stfisccap').^[The omitted, baseline category in the regression model captures weak states with little domestic access to revenue.] The model also includes several predictors from the previous analysis: lagged judicial independence, lagged GDP growth rate, whether it is a presidential election, the degree of urbanization, and lagged physical integrity and its square term.  


**Stage 2 predictors**



## Analysis 1 Results




```{r stage2-models, echo=FALSE}
###lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  + frefair.1lag  + osorg.1lag   +       + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + n.events.pre  +  unscheduled_election + regional_success_lag



model.fraud_ps <- lm(v2elirreg.inv ~ propensity_score , data = nelda2.sub.fraudpred)

model.fraud.covars <- lm(v2elirreg.inv ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag     + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + n.events.pre + unscheduled_election + regional_success_lag , data = nelda2.sub.fraudpred)
#summary(model.fraud.covars)

model.fraud.all <- lm(v2elirreg.inv ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag     + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + n.events.pre + unscheduled_election + regional_success_lag + propensity_score, data = nelda2.sub.fraudpred)
#summary(model.fraud.all)

model.fraud.resid <- lm(.resid ~ n.events.pre +  regional_success_lag , data = nelda2.sub.fraudpred)


models5 <- list()
models5[['Model 3']] <- model.fraud_ps
models5[['Model 4']] <- model.fraud.covars
models5[['Model 5']] <- model.fraud.all
models5[['Model 6']] <- model.fraud.resid





modelsummary::modelsummary(models5, gof_omit = "BIC|Log", longtable=F, stars = T, title = 'Stage 2 models of election fraud (OLS)', coef_map = 
               c("(Intercept)" = "(Intercept)",  "lji.lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", "v2elirreg.inv" = "Election fraud",
                 "winner.share.upd" = "Incumbent vote-share",
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                  "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "n.events.massmob.pre" = "Number of pre-election protests", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity",
                 "state.owned.1lag" = "Private ownership (lag)",
                 "patronage.appts.1lag" = "Professional bureaucracy (lag)",
                 "v2elembaut" = "EMB autonomy",
                 ".fitted" = "Expected fraud level",
                 "propensity_score" = "Protest risk"
                 ))
                 

```






Table \@ref(tab:stage2-models) shows the results of the second-stage models from both approaches. Model 3 shows that an increase in the pre-election propensity score of protest is associated with a sizable and statistically significant *increase* in subsequent manipulation. Model 4 shows that a recent successful protest movement in the region is also associated with more severe election fraud, while controlling for a variety of possible confounders. Model 5 includes all controls, the regional protest dummy, and the protest propensity score; the regional protest dummy remains positive and significant at the conventional level. Finally, Model 5 shows that increasing pre-election protests and successful electoral protest in the region are positively associated with increased fraud *relative to each country's baseline*. In other words, increasing protest risk is not only associated with higher levels of fraud cross-nationally, but also with intensifying fraud *within countries*.

Where the regime-weakness model predicts that protest risk should deter manipulation, the results find just the opposite: incumbents appear to manipulate more heavily under riskier conditions. This is a theoretically significant result, and is also evidence in favor of the signaling mechanism described in the theory section; as the risk of post-election protest rises, incumbents are incentivized to manipulate more heavily to signal their resources and capacities to within-system elites. 


```{r, include=F, echo=FALSE, message=FALSE, warning=F, eval=FALSE}
## Anova
anova_dat <- nelda2.sub %>% dplyr::select(inc.replaced.prot, v2elirreg.inv,  jucon.1lag, legcon.1lag, gdpgro.1lag,  osorg.1lag, natlpartyorg.lag,  urban.pct, n.events.pre, state.owned.1lag, patronage.appts.1lag, v2elembaut)

# boxplots and t-tests for the 4 variables at once
for (i in 2:ncol(anova_dat)) { # variables to compare are variables 2 to end
  boxplot(anova_dat[, i] ~ anova_dat$inc.replaced.prot, # draw boxplots by group
    ylab = names(anova_dat[i]), # rename y-axis with variable's name
    xlab = "Incumbent replaced after protest"
  )
  print(paste(c("ANOVA for variable ", (names(anova_dat[i])))))
  print(t.test(anova_dat[, i] ~ anova_dat$inc.replaced.prot)) # print results of t-test
}

aov_output <- aov(state.owned.1lag ~ inc.replaced.prot, data = nelda2.sub)
summary(aov_output)
```


```{r, echo=FALSE, message=FALSE, include=FALSE, warning=FALSE}
#model.loss_ps <- glm(inc.lose ~ propensity_score + .fitted , data = nelda2.sub.fraudpred, family = binomial(link="logit"))

#model.lose_reg <- glm(inc.lose ~ .fitted + regional_success_lag , data = nelda2.sub.fraudpred, family = binomial(link="logit"))
#summary(model.lose_reg)

#model.lose_events <- glm(inc.lose ~ .fitted + n.events.pre , data = nelda2.sub.fraudpred, family = binomial(link="logit"))
#summary(model.lose_events)


#summary(model.fraud.resid)

#model.lose_resid <- glm(inc.lose ~  .resid  , data = nelda2.sub.fraudpred, family = binomial(link="logit"))
#summary(model.lose_resid)

model.protest_resid <- glm(any.ecav.protest ~  .resid  , data = nelda2.sub.fraudpred, family = binomial(link="logit"))
#summary(model.protest_resid)
```




# Analysis 2: Severity of manipulation and protest duration


**Data and modeling strategy** In Analysis 2, I test whether more extensive electoral manipulation is associated with greater incumbent strength against post-election challenges. The main variables used in this analysis are a count measure of *protest duration* in days, taken from the ECAV dataset. Once protests begin (whether instrumental or multi-purpose), authoritarian incumbents have strong incentives to defeat a public challenge to their legitimacy, while committed opponents have similar incentives to stand firm. Both should be expected to put significant effort into defeating / maintaining the protest movements, as a general rule, and the process by which the challenge is settled helps reveal previously hidden information about each side; it is possible to observe whether incumbents who produce extensive election fraud are stronger or weaker than those that do not.

ECAV data is in event-day format; to capture duration, I take the number of days between the last protest recorded in ECAV and the first, for each election.^[The duration of a protest wave with only one protest is taken to be one day.] Protest wave duration is modeled using Cox proportional hazards models, which have the following formula:

$$h(t | X)=h_0(t)* exp(\beta_fX_f+\beta_vX_v+ ... +\beta_pXp_)$$
The formula models the risk that a protest movement will fail at time $t$ given covariates $X$ is a function of the baseline hazard function, the degree of electoral fraud ($X_f$), the incumbent's vote-share ($X_v$) and a set of control variables. For the survival models, protest movements that were ultimately defeated by the incumbent are considered 'dead' after the last day of protest. Protest movements that prevailed in unseating the incumbent are included, and are marked as 'alive.' One threat to inference in this design derives from the 90-day post-election window considered by ECAV. It is possible that some protest movements considered 'dead' are, in fact, 'alive,' if subsequent protests occurred after 90 days and are unobserved. To address this, observations are assumed to be right-censored if a protest occurred within three-weeks of the 90-day cutoff in a subset of models.


The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0) %>% nrow()` of the `r nelda2.sub %>% nrow()` elections, no protests are recorded. Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest duration. As the figure shows, while election fraud is common among this set of regimes, protests are rare, and, when they do occur, are usually short in duration. These distributions\textemdash extensive election fraud combined with large numbers of non-protest\textemdash provides some early indication that the regime-weakness model has limited support. Still, approximately one-quarter of elections experienced some protest, and so protest initiation is modeled using standard logistic regression. Protest wave duration is modeled using Cox proportional hazards models.



```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}

p.hist.events <- ggplot(nelda2.sub, aes(x=n.events.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()

p.hist.duration <- ggplot(nelda2.sub, aes(x=d.events.num)) + geom_histogram(bins=100) + xlab("Protest wave duration") + theme_bw()


p.hist.fraud <- ggplot(nelda2.sub, aes(x = v2elirreg.inv)) + geom_histogram(bins=100) + xlab("Election fraud") + theme_bw()


 ggplot(nelda2.sub, aes(y = v2elirreg.inv, x= jitter(n.events.post))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("n.events.post (jittered)") + theme_bw()
 
#  ggplot(nelda2.sub, aes(y = v2elirreg.inv, x= max_size)) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("max size") + theme_bw()
```


```{r elex-hist, fig.cap="Histograms of election fraud and protest duration", fig.height=4, echo=FALSE, message=FALSE, warning = FALSE}
plot.combined.hist <- cowplot::plot_grid(p.hist.fraud, p.hist.duration, ncol=1)
plot.combined.hist
```

**Explanatory variables**: 


In addition to *election fraud*, I include a measure of the *incumbent party's vote-share* taken from the V-Dem dataset.^[This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia, where links to official sources were available, and Psephos.] The original V-Dem variable indicates the *winner's* vote-share. However, I exclude all cases where the incumbent lost the election outright. This limits the observations to those that fit the population of interest: protest following an election in which the incumbent claimed victory, possibly through reliance on election fraud; it also ensures that the vote-share variable captures the incumbent's position. 

Excluding cases where incumbents stepped down after losing an election avoids introducing measurement error in the dependent variable, due to a quirk of the ECAV dataset, which codes the 'side' of the protest (pro- or anti-government) as missing during the period between an election and the swearing in of a new government in cases where the incumbent loses. Including these cases thus runs the risk of conflating pro- and anti-incumbent protest, along with cases where protesters goals are genuinely unknown. However, excluding them also influences the interpretation of the results from these models. Threats to inference from this section are evaluated in more detail in the discussion section. 


The need to exclude cases where incumbents lost the election introduces the possibility of selection bias. Most crucially, it could be the case---as deterrence models predict---that incumbents who believe they will face post-election protest choose not to manipulate (in effect, voluntarily choosing to lose the election). Only the strongest incumbents would then 'select in' to the risk of protest by manipulating extensively. While, as with all observational studies, it is possible that unknown or unobservable variables may introduce bias, I take several steps to reduce this risk. 

First, the results of Analysis 1 show that increased risk of protest leads incumbents to *intensify* their manipulation efforts, rather than diminish them. This is the opposite of what we would expect to see if a selection effect were at work. Rather than incumbents selecting out of manipulation when they face higher risks, they appear to commit to more extensive manipulation on average. Second, in the survival models below, I control for measures of opposition strength and grievances, to help isolate the relationship between manipulation severity and protest duration. Finally, I include regression models of incumbent loss in the appendix. As in Analysis 1, these tests show that ... [COME BACK AFTER ADDING TO APPENDIX] 


**Control variables** The theory suggests that increasing election manipulation is indicative of regime strength, but an actor's strength---like its close analogue, power [@dahl_concept_1957]---can only be measured relative to that of another. An incumbent who produces a high level of election manipulation may be 'strong' when facing a fragmented and unpopular opposition, but would be less so if the opposition were more formidable. As a result, it is necessary to control for potential indicators of opposition strength.

First, I include the *number of pre-election protests* counted in the ECAV dataset, which is likely to be an important possible confounder; more numerous pre-election protest may affect the level of election fraud (as the deterrence logic predicts) as well as being directly tied to the likelihood of post-election protests. Next, I control for the *median size of protest*, calculated using ECAV data.^[ECAV data codes protest size as a categorical variable, ranging from 1 (< 10 participants) to 5 (> 10,000). I take the median size category of all the events in the protest wave as a measure of average protest size.] This is important as a control for opposition strength, as well as for the possibility that incumbents might tolerate smaller protests. Since non-democratic governments may be able to adjust the electoral calendar strategically, to disrupt opposition coordination [@nygardTimingMattersImpact2020], I include a variable from NELDA that indicates *off-schedule elections*. 

Since higher repressiveness may weaken *or* intensify opposition [@chenowethStateRepressionNonviolent2017], I include a lagged measure of *physical integrity* index from V-Dem (v2x_clphy), and its square term. Indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), *legislative constraints* on the executive (V-Dem variable 'v2xlg_legcon', and the availability of *alternative sources of information* ('v2x_freexp_altinf' from V-Dem) are also employed as straightforward indicators of opposition strength. Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019]. 

Other controls are included for factors that might separately cause variation in electoral manipulation and protest duration. First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@simpser2013] and a greater risk of protest. Another NELDA variable indicates if the incumbent executive is not running due to *term limits*, which may produce elite fragmentation [@hale_patronal_2014] that affects both protest risk and electoral manipulation. De facto *judicial independence* [@Linzer2015] is included, since judicial independence can reduce manipulation [@Harvey2022] and protest risk [@chernykh2014]. The *lagged GDP growth rate*, taken from V-Dem, is used to capture economic grievances: lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance (increasing protest risk, but potentially deterring manipulation under the regime-weakness framework).

In addition, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby due to opposition learning [@Bunce2010;@Beissinger_2007]. Such a dynamic could plausibly increase the severity of protest in a nearby country, while also reducing the incumbent's use of fraud. I control for this using each country's geopolitical region as coded in Teorell et al [-@qog_teorell], in conjunction with binary variables indicating a successful protest movement from NELDA. A 'successful protest' is recorded if a protest movement resulted in either a new election being run or the incumbent stepping down. The *regional diffusion* dummy variable takes on a value of one for all countries in the same region as a successful protest, for elections within one calendar year of the successfully overturned election.^[For example, the variable takes on a positive value for elections held in the post-communist region, within one year of the September 24, 2000 election in Serbia that sparked the 'Bulldozer Revolution'. All told there are 14 successful electoral protests in the dataset: Bulgaria 1990, Burkina Faso 1997, Lesotho 1998, Cote d'Ivoire 2000 and 2010, Peru 2000, Serbia 2000, Mali 2002, Georgia 2003, Ukraine 2004, Azerbaijan 2005, Kyrgyzstan 2005, Haiti 2006, and Cameroon 2007.]    

<!-- I include dummy variables that provide a categorical evaluation of state capacity. These are taken from the ordinal version of the V-Dem variable "state source of fiscal revenue" (v2stfisccap). If the V-Dem data codes a state as not capable of raising revenue or primarily relying on loans and foreign aid, I code it as a *weak state*. If the state is described as primarily relying on resource rents and expropriation, I code it is a *rentier state*. States that rely primarily on taxes on property or economic transactions are coded as *tax states*. This distinction is relevant since access to natural resource rents may both make protest less likely and repression more effective [@Girod2018].   -->


Summary statistics for all of these variables are shown in the appendix. To address concerns about multicollinearity, I also present models in the appendix that exclude control variables associated with pre-election protest.






```{r echo=FALSE, message=FALSE, warning=FALSE}
### Creating dataframe with fitted values for fraud

m.fraud.pred_base <- lm(v2elirreg.inv ~ n.events.pre + regional_success_lag, data = nelda2.sub)

m.fraud.pred <- lm(v2elirreg.inv ~ jucon.1lag +  gdpgro.1lag + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag  + v2elembaut, data = nelda2.sub)

#summary(m.fraud.pred)


```



```{r, include = F, echo=FALSE}
### Calculate propensity scores for protest, but use lagged electoral fairness as variable

complete_cases <- complete.cases(nelda2.sub.fraudpred[, c("any.ecav.protest", "lji.lag" , "legcon.1lag" , "gdpgro.1lag" , "altinf.1lag"  , "frefair.1lag"  , "osorg.1lag"   ,   "presidential" , "inc.termlimit.num"  ,  "physinteg.1lag" , "physintlag.sq" , "urban.pct" , "n.events.pre" , "unscheduled_election" , "regional_success_lag", ".fitted")])

model.protest.null <- glm(any.ecav.protest ~ 1, family = binomial(link = "logit"), data = nelda2.sub.fraudpred[complete_cases, ])
summary(model.protest.null)

model.protest.nofraud <- glm(any.ecav.protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  + frefair.1lag  + osorg.1lag   +       + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + n.events.pre  +  unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub.fraudpred[complete_cases, ])
#summary(model.protest.nofraud)

nelda2.sub.fraudpred$propensity_score <- NA

# Assign scores only to complete cases
nelda2.sub.fraudpred$propensity_score[complete_cases] <- predict(model.protest.nofraud, type = "response")

propensity_scores <- predict(model.protest.nofraud, type = "response")

```


 
```{r binarytab, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
count_table <- table(cut(nelda2.sub$winner.share.upd, breaks = c(0, .45, .55, 1), 
    include.lowest = TRUE, labels = c("Low", "Medium", "High")),
      cut(nelda2.sub$v2elirreg.inv, breaks =quantile(nelda2.sub$v2elirreg.inv, probs = c(0, 1/3, 2/3, 1), na.rm=T), 
    include.lowest = TRUE, labels = c("Low", "Medium", "High")),
      nelda2.sub$any.ecav.protest)



prop_count_table <- nelda2.sub %>% filter(inc.lose == 0) %>%
  mutate(winnershare_bin = cut(winner.share.upd,  breaks = c(0, .45, .55, 1), 
    include.lowest = TRUE, labels = c("Low (0-45%)", "Medium (45-55%)", "High (55-100%)")),
         fraud_bin = cut(v2elirreg.inv,  breaks =quantile(nelda2.sub$v2elirreg.inv, probs = c(0, 1/3, 2/3, 1), na.rm=T), 
    include.lowest = TRUE, labels = c("Low", "Medium", "High"))) %>%
  group_by(winnershare_bin, fraud_bin) %>%
  summarise(pct_positive = mean(any.ecav.protest) * 100)

prop_count_table <- pivot_wider(prop_count_table, names_from = fraud_bin, values_from = pct_positive)
prop_count_table <- prop_count_table[1:3, 1:4]
prop_count_table <- round(prop_count_table[,2:4], 2)
prop_count_table <- cbind('Incumbent vote-share' = c("Low", "Medium", "High"), prop_count_table)




prop_count_table %>%
  kable(format = "latex", booktabs = TRUE, caption = "Percentage of cases in which post-electoral protests occurred") %>%
  add_header_above(c(" " = 1, "Electoral fraud" = 3)) %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

<!-- As a plausibility check, I first examine patterns of protest initiation. The regime-weakness model expects initiation to increase with fraud, holding vote-share constant; while the regime-strength model expects a less consistent pattern. Table \@ref(tab:binarytab) categorizes elections into 3 x 3 bins based on incumbent vote-share and degree of electoral fraud, and reports the percentage of elections in each category that were followed by protest.  For this table, vote-share is considered low if it less than 45%, medium between 45% and 55%, and high if greater than 55%. The middle category is intended to capture a 'close' election. Election fraud is divided into terciles. These categories can be thought of as relatively clean (*election fraud* is between -2.33 and 0.5), moderately manipulated (0.5 to 1.46) and heavily manipulated (1.46 to 3.2). Two patterns appear in the data. First, protest initiation is most likely when incumbent vote-share is low and fraud is high. This is expected under both models. For the regime-weakness model, this is taken as a sign that opposition actors launch protests instrumentally against the weakest (i.e. least popular) incumbents. For the regime-strength model, this pattern could emerge as electoral manipulation focuses mass grievances. However, we do not observe a consistent increase in protest likelihood as manipulation increases, as the regime-weakness model would predict. Instead, we see relatively consistent rates of protest initiation at medium vote-share across levels of manipulation. And, at high incumbent vote-shares, there is even a decrease in protest initiation as manipulation increases from moderate to high. This comports with the mixed evidence found in prior research on protest initiation and manipulation, and is evidence that the regime-strength model is on a plausible foundation (specifically, that protests can occur even when the incumbent is predicted to be 'strong' under either model). Consider the cell where vote-share is high and election fraud is low---this is the strongest (i.e. clearly popular) type of regime in the regime-weakness model. Nevertheless, more than a quarter of such cases are protested. This is not so different from the cell which is most likely to contain strong regimes in the regime-strength model (high vote-share, high manipulation), in which thirty-two percent of cases launch protest. While not dispositive, this result indicates that the regime-strength model and regime-weakness models are equally compatible with patterns of protest initiation, as expected.

-->

```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}

p.hist.events <- ggplot(nelda2.sub, aes(x=n.events.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()

p.hist.duration <- ggplot(nelda2.sub, aes(x=d.events.num)) + geom_histogram(bins=100) + xlab("Protest wave duration") + theme_bw()


p.hist.fraud <- ggplot(nelda2.sub, aes(x = v2elirreg.inv)) + geom_histogram(bins=100) + xlab("Election fraud") + theme_bw()

nelda2.sub <- nelda2.sub %>% mutate(fraud_transform = v2elirreg.inv - min(v2elirreg.inv, na.rm=T) + 1)
nelda2.sub <- nelda2.sub %>% mutate(vs_fraud_ratio = winner.share.upd / fraud_transform)

# ggplot(nelda2.sub, aes(x = vs_fraud_ratio, y= jitter(any.ecav.protest))) + geom_point(aes(color = v2elirreg.inv)) + xlab("Ratio") + theme_bw() + scale_color_viridis_b()
 
#  ggplot(nelda2.sub, aes(x = v2elirreg.inv, y= jitter(any.ecav.protest))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("Fraud") + theme_bw()
  
#   ggplot(nelda2.sub, aes(x = v2elirreg.inv, y= n.events.post)) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("Fraud") + theme_bw()
 
#  ggplot(nelda2.sub, aes(x = winner.share.upd, y= jitter(any.ecav.protest))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("Winner share") + theme_bw()
  
  
#    ggplot(nelda2.sub, aes(x = winner.share.upd, y= n.events.post)) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("Winner share") + theme_bw()

  
  
#   ggplot(nelda2.sub, aes(x = vs_fraud_ratio, y= d.events.num)) + geom_point(aes(color = v2elirreg.inv)) + xlab("Election fraud") + theme_bw() + scale_color_viridis_b()
 
  ggplot(nelda2.sub, aes(x = v2elirreg.inv, y= d.events.num)) + geom_point(aes(color = winner.share.upd)) + scale_color_continuous_divergingx(palette = "ArmyRose", mid = .5) +  xlab("Fraud") + theme_bw()
 
 
#  ggplot(nelda2.sub, aes(y = v2elirreg.inv, x= max_size)) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("max size") + theme_bw()
  library(colorspace)
  
# ggplot(nelda2.sub %>% filter(inc.lose != 1) , aes(x=winner.share.upd, y=n.events.post, color = v2elirreg.inv)) +
#    geom_point() +
#    scale_color_continuous_divergingx(palette = "ArmyRose", n_interp = 11, cmax1 = 500, cmax2 = 500) +
#    theme_bw() +
#   coord_cartesian(ylim = c(0, 100))
 
 
   
# ggplot(nelda2.sub %>% filter(inc.lose != 1), aes(x=winner.share.upd, y=jitter(any.ecav.protest), color = v2elirreg.inv)) +
#    geom_point() +
#    scale_color_continuous_divergingx(palette = "ArmyRose", n_interp = 11, cmax1 = 500, cmax2 = 500) +
#    theme_bw()
 
 
#  ggplot(nelda2.sub %>% filter(inc.lose != 1 & d.events.num > 0), aes(x=winner.share.upd, y=d.events.num, color = v2elirreg.inv)) +
#    geom_point(aes(shape = as.factor(protest.success))) +
#    scale_color_continuous_divergingx(palette = "ArmyRose", n_interp = 11, cmax1 = 500, cmax2 = 500) +
#    theme_bw()
  

  library(latticeExtra)
  # showing data points on the same color scale 
levelplot(any.ecav.protest ~ winner.share.upd * v2elirreg.inv, nelda2.sub %>% filter( protest.success == 0 & inc.lose == 0), 
          panel = panel.levelplot.points, cex = 1.2
    ) + 
    layer_(panel.2dsmoother(..., n = 300))


  # showing data points on the same color scale 
levelplot(d.events.num ~ winner.share.upd * v2elirreg.inv, nelda2.sub %>% filter(d.events.num > 0 & protest.success == 0 & inc.lose == 0), 
          panel = panel.levelplot.points, cex = 1.2
    ) + 
    layer_(panel.2dsmoother(..., n = 300))


#  testlogit <- glm(protest.success ~ v2elirreg.inv*winner.share.upd + presidential + inc.termlimit.num + n.events.pre, data = nelda2.sub)
#summary(testlogit)
#sjPlot::plot_model(testlogit, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.08, 1.96]"))
```


## Analysis 2 Results




```{r, include=FALSE}
library(survival)
library(ggsurvfit)




nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(inc.replaced.prot == 1, 0, surv_status))
nelda2.sub <- nelda2.sub %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
                                                                       ifelse(d.events.num >= 69, 0, 1)))




cox_base <- coxph(Surv(d.events.num, surv_status) ~   winner.share.upd*v2elirreg.inv + presidential  + n.events.post , data = nelda2.sub %>% filter(inc.lose != "1" &  n.events.post > 0))
summary(cox_base)




cox1 <- coxph(Surv(d.events.num, surv_status) ~  lji.lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct  +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose != "1" &  n.events.post > 0))
summary(cox1)



cox1.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~ lji.lag + legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct  +  unscheduled_election + regional_success_lag + v2elirreg.inv   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0))
summary(cox1.cen)





#cox1t2 <- coxph(Surv(d.events.num, surv_status) ~ legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election +  regional_success_lag + n.events.post + v2elirreg.inv + tt(n.events.post), data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), tt = function(x, t, ...)  x* nsk(t, knots=c(10, 70, 90),
#Boundary.knots = FALSE))
#summary(cox1t2)









cox2 <- coxph(Surv(d.events.num, surv_status) ~ lji.lag + legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag +  + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +  unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
summary(cox2)
p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.29, 2.2]")) #This is the 10th and 90th quantiles
p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()


cox2.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~ lji.lag +  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct  + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
summary(cox2.cen)


cox3.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~ lji.lag +  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag +  v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + unscheduled_election + regional_success_lag + n.events.post + median.participants, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
summary(cox3.cen)



```





Table \@ref(tab:cox-models-tab) reports the results of Cox proportional hazard models of protest duration. In these models, positive coefficients indicate factors that are associated with shorter-lived protests; those with a negative sign indicate factors associated longer-lasting protest. 

Model 7 is a baseline with minimal controls. Models 8 and 10 do not include interactions, while the remaining models interact *election fraud* with incumbent vote-share. One threat to inference in this design derives from the 90-day post-election window considered by ECAV. Models 10, 11, and 12 incorporate the possibility of right-censoring in cases where the last observed protest occurred in the last three weeks of the 90-day observation window. This amounts to 15 of the observations in the sample. Model 12 is the most restrictive model, which includes right censoring as well as a control variable for the median number of participants during the protest wave. There is considerable missingness in this variable, resulting in a loss of roughly one-third of the sample. 

The coefficient for *election fraud* is positive and significant in Models 7 and 8, indicating that more fraudulent elections are associated with quicker government victories. Exponentiation of the coefficient for election fraud gives a hazard ratio of 1.3, indicating that for each one-unit increase in fraud, a protest movement becomes thirty percent more likely to end at any given time. This is inconsistent with the regime-weakness model, but consistent with Hypothesis [FILL IN]. 

In Model 9, incumbent vote-share is interacted with *election fraud*. Figure \@ref(fig:margin-dur-fig) shows the predicted hazard ratios. As shown in the figure, for incumbent vote-shares near and below fifty percent, large-scale fraud significantly increases the likelihood that a protest movement will end. This result is in line with Hypothesis [FILL IN] and the regime-strength model, and shows that only as vote-share increases well above fifty percent---creating a stronger signal of popularity---does a government in a low-fraud election have the same likelihood of defeating a protest movement at a given time. 

The results of the baseline models are consistent with those models which incorporate censoring. The coefficient for *election fraud* in Models 10 and 11 retains the same approximate size, sign, and level of significance as in the corresponding models without censoring. This suggests that the main models are not drawing biased inferences through censoring of longer-lasting protest movements. Model 12, adds the median number of participants at events in the protest wave as a control for the potential perceived threat of the protest movement by the incumbent. Including this control variable reduces the sample size due to missingness, and standard errors increase as a result. Nevertheless, the coefficients on election fraud and the interaction term retain their size and direction.



```{r margin-dur-fig, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
p.duration.margin
```



The relationship between the level of fraud and the survival of protest movements can also be demonstrated non-parametrically using Kaplan-Meier curves for varying levels of fraud. In Figure \@ref(fig:km-fig), observations are divided into three groups. High fraud observations are those where *election fraud* is one standard deviation above the mean or more, low fraud observations are those one standard deviation below the mean or less, with moderate levels in between. The figure shows a clear difference between the low-fraud group and the other two. Low-fraud cases are much more likely to see their protest movements survive past the first day than their counterparts. For much of the range of the data, they remain roughly twice as likely to survive at any given date than protest movements in medium- or high-fraud cases. This is evidence in favor of the idea that high levels of fraud indicate stronger regimes. 

```{r km-fig, fig.cap="Kaplan-Meier curves for protest duration across levels of fraud", fig.height= 5, echo=F, message=F, fig.pos='H'}
nelda2.sub <- nelda2.sub %>% mutate(fraud_level = ifelse(v2elirreg.inv >= (.98 + .97), "High",
                                                            ifelse(v2elirreg.inv <= (.98 - .97), "Low", "Moderate")))
nelda2.sub.km <- nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0)


p.km <- survfit2(Surv(d.events.num, surv_status) ~ fraud_level, data = nelda2.sub.km %>% filter(d.events.num < 300)) %>% 
  ggsurvfit(linetype_aes = T) +
#   add_confidence_interval() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  )

p.km
```




```{r cox-models-tab, echo=FALSE, warning=FALSE}
models3 <- list()
models3[['Model 7']] <- cox_base
models3[['Model 8']] <- cox1
models3[['Model 9']] <- cox2
models3[['Model 10']] <- cox1.cen
models3[['Model 11']] <- cox2.cen
models3[['Model 12']] <- cox3.cen





modelsummary::modelsummary(models3, fmt=2, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Table 4: Cox proportional hazard models of protest duration', coef_map = 
               c("(Intercept)" = "(Intercept)", "lji.lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.post" = "N. of post-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", 
                 "median.participants" = "Median protest size",
                                  "winner.share.upd" = "Incumbent vote-share",
                 "v2elirreg.inv" = "Election fraud",
"v2elirreg.inv:winner.share.upd" = "Election fraud:Inc. vote-share"
                 ))
```



# Discussion

Analysis 1 tests how electoral manipulation responds to increases in protest risk cross-nationally, and over time within countries. Contrary to conventional expectations, the results consistently demonstrate that election fraud *increases* with protest risk. Analysis 2, on the duration of protest events, points in the same direction: more election manipulation is associated with quicker incumbent victories over protest movements. In other words, in real contests in which opposition and incumbent strengths are tested against each other, better manipulators reveal themselves to be stronger rather than weaker. 

The contributions of this study are twofold. First, it directly contradicts a central implication of the regime-weakness model of manipulation. As described earlier in this paper, a sizable literature is founded on the assumption that increasing protest risk generally reduces incumbents' incentives to manipulate (for fear of appearing weak), leading to cleaner elections. Here we observe the opposite relationship. Second, this study investigates what happens when we depart from the equilibrium path predicted in the regime-weakness model. When electoral manipulation is high and post-election protests do occur, the purported deterrent effect of protest risk has failed. If electoral manipulation signals regime weakness, we should observe that incumbents who manipulate more severely have a harder time surviving protest movements. Once again, we find the opposite relationship.

Both findings are incompatible with the 'regime-weakness' model of manipulation, but are fully consistent with a model in which electoral manipulation is a costly signal of strength that can induce elite loyalty during challenges. Further, these relationships hold despite controlling for indicators of other possible explanations of the relationship between low fraud and long-lived protest, including civil society strength, media openness, regional electoral protest waves, number of protests and median protest size. They are also robust to different model specifications and operationalizations of key variables. While this study does not purport to establish perfect causal identification, these robust correlations provide substantial empirical support for the regime-strength model of manipulation.

This view of election manipulation also helps us make sense of a secular pattern that is potentially puzzling under the regime-weakness framework: if extensive manipulation signals incumbent weakness, why do so many incumbents manufacture their way to highly implausible margins of victory? Examples abound: Paul Kagame in Rwanda's 2017 election (98.8 percent vote-share), Islam Karimov in Uzbekistan's 2015 election (91.8 percent),  Alexander Lukashenko in the 2020 Belorussian election (a comparatively modest 81 percent) are illustrative. These results are explicable in the regime-strength model, which reasons that they credibly demonstrate incumbents control over resources and low-level agents, which in turn helps induce elite loyalty.


In rejoinder, it might be asked: if election manipulation credibly signals strength, why do protests occur after highly manipulated elections? In other words, why do we not observe a reverse deterrent effect whereby extensive manipulation deters strategic protest? First, the frequency of severe manipulation and the rarity of protest shown in Figure \@ref(fig:elex-hist) suggests that such an effect may occur in some cases; strategic protest entrepreneurs may in fact shy away from initiating protests if they perceive the incumbent to be well-resourced and capable of mobilizing large numbers of agents. Second, and more broadly, instrumental calculations about overthrowing incumbents are but one component of a long and varied literature on the causes of protest, which has identified multiple other expressive and strategic motives. These other motives may foster protest movements in the aftermath of rigged elections, even when severe electoral manipulation indicates considerable regime strength and a low likelihood of unseating the incumbent. 





<!--
Additionally, it is likely that when strategic protest entrepreneurs are considering initiating protest, they must consider how likely it is that ordinary citizens will join them [@buenodemesquitaRegimeChangeRevolutionary2010]. In the absence of large-scale election manipulation to serve as a coordination device [@kuntz_more_2009;@tucker2007], they may judge that protests will be unsuccessful despite the regime's apparent low popularity. As Schedler writes, "Grievances provide moral justifications for collective protest: we need to do something. Opportunities provide strategic justifications: we can achieve something" [@Schedler2013,p.304]. In essence, they may be gambling that, under pressure from mass protests and facing low genuine popularity, elites and state officials will split from the incumbent. In other words, they may hope that by initiating protest when the incumbent has demonstrated low popularity, the cooperation between principals and agents demonstrated by effective fraud will begin to fray [@Hale2019;@satoEliteCoordinationPopular2019]. This kind of gamble may be especially sensible in more deeply authoritarian contexts, where information on incumbent strengths and weaknesses is scarce [@Pop-Eleches2015] and preference falsification more widespread [@Kuran1991]. Unfortunately, the duration analysis suggests this is often a losing bet, since it is exactly these scenarios wherein post-election protest is most likely to fail quickly. Untangling the relationship between strategic and non-strategic protest when election manipulation indicates sources of strength is an important direction for future research.

A very brief consideration of the cases of successful electoral protest in comparison with the failed cases can help aid interpretation of the results. Table 3 shows the mean values for several key variables for two groups: cases where protests resulted in incumbents' ouster, and cases where protest movements were quashed. The table also shows p-values for t-tests comparing the means of the two groups. While the level of fraud is typically higher in the successful protest group than the unsuccessful group, this difference is not significant. By contrast, incumbent vote-shares are lower in the successful-protest and protest groups than their counterparts (though this difference is only significant at the p < .1 level). Only two covariates are different at the conventional level of significance. Cases where protest succeeds in topping governments tend to be more heavily urbanized (by 10 percentage points) and have an economy less dominated by the state (by sixty-four percentage points!). This is further evidence that a high degree of electoral manipulation is not indicative of regime weakness; more structural variables appear to drive regime strength, in keeping with the model outlined here.

```{r, include=F, warning=FALSE, message=FALSE, echo=FALSE}
## Anova for protest success
anova_dat <- protest.cases %>% dplyr::select(protest.success, v2elirreg.inv,  jucon.1lag, legcon.1lag, gdpgro.1lag,  osorg.1lag, natlpartyorg.lag,  urban.pct, n.events.pre, state.owned.1lag, patronage.appts.1lag, v2elembaut, winner.share.upd)

# boxplots and t-tests for the 4 variables at once
for (i in 2:ncol(anova_dat)) { # variables to compare are variables 2 to end
  boxplot(anova_dat[, i] ~ anova_dat$protest.success, # draw boxplots by group
    ylab = names(anova_dat[i]), # rename y-axis with variable's name
    xlab = "Incumbent loss"
  )
  print(paste(c("ANOVA for variable ", (names(anova_dat[i])))))
  print(t.test(anova_dat[, i] ~ anova_dat$protest.success)) # print results of t-test
}

aov_output <- aov(state.owned.1lag ~ inc.lose, data = nelda2.sub)
summary(aov_output)
```

```{r anova-table, echo=FALSE, warning=FALSE, message=FALSE}
#p.values <- c(3.009e-15, 6.014e-11, 6.149e-10, 0.01146,  3.758e-07,  0.6281, 0.003519,
#              0.1852, 0.007405, 0.003086, 2.2e-16)
#win_lose_diff <- c(.87, -0.15, -0.16, .06, -.47, .04, 5.3, -1.46, -.265, -.257, -.893)
#variables <- c("Election fraud", "Judicial constraints", "Legislative constraints", "GDP growth (lag)", "Civil society openness (lag)", "Natl. party institutionalization (lag)", "Urban",
#               "N. pre-election protests", "Privatization", "Professional bureaucracy", "EMB autonomy")
#anova_dat <- data.frame(cbind(variables,win_lose_diff, p.values))
#anova_dat <- anova_dat %>% mutate(win_lose_diff = as.numeric(win_lose_diff))
#anova_dat <- anova_dat %>% mutate(p.values = as.numeric(p.values))
#anova_dat <- anova_dat %>% rename(Variables = variables)
#anova_dat <- anova_dat %>% rename("Win-Loss Difference" = win_lose_diff)
#anova_dat <- anova_dat %>% rename("P-value" = p.values)

##Commented out data above is for election win / loss by incumbent
# Below we have protest success / failure

anova_dat <- read.csv(here::here("Data", "anova_protest_success.csv"))

anova_dat <- anova_dat %>% rename("Success-Fail Difference" = Success.failure.diff.)
anova_dat <- anova_dat %>% rename("P-value" = p.value)
anova_dat <- anova_dat %>% rename("Protest Failure" = Protest.failure)
anova_dat <- anova_dat %>% rename("Protest success" = Protest.success)


protest_table <- flextable(anova_dat)
protest_table <-  set_caption (protest_table, "Table 4: Mean variable values for failed and successful protests, with ANOVA p-values") 
protest_table

```



The causes of the post-communist cases, which account for most of the seven cases, has been much debated [@LucanWay2008;@Bunce2009]. Yet for most of these cases, election fraud either held steady after the 'successful' revolution, or increased before partially declining again some years later. These cycles, identified as a feature of 'patronal politics' [@hale_patronal_2014] could indicate that in these cases, fraud patterns and protest were caused by more fragmented elites who failed to resolve the principal-agent problems needed to engage in mass protest and control the streets. A thorough qualitative comparative analysis of the cases would be informative. 
-->


What does all this tell us? First, these findings support and extend the line of research initiated by Simpser [@simpser2013]. Grounding Simpser's approach in the literature on the principal-agent dynamics of manipulation [@Harvey2019;@rundlett2016deliver;@Harvey2022] deepens our understanding of why election manipulation is a costly signal (incumbents must overcome severe principal-agent problems) and shows why manipulation need not be excessive or blatant to induce elite loyalty (manipulation efforts are intelligible to elites embedded within the network of principals and agents). Additionally, this study illustrates how the mechanisms which underpin Simpser's model of manipulation-as-strength---coordination effects among pro-regime agents and costly signaling of incumbent resources---can function even when such signals of strength have failed to 'deter' protest.

Second, these results may inform how researchers think about the causes of election manipulation. They suggest that the degree to which the opposition's capacity to organize street protest can limit election fraud has been overestimated. Consequently, theories of electoral manipulation that emphasize alternatives---such as resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022]---should be further developed, and may yield new policy interventions. In particular, it may change some current understandings of the role of election monitors, courts, and other third parties in upholding election integrity. As noted earlier, prior studies argue that these actors restrain manipulation by revealing information about fraud and intensifying protest risk. However, if more extensive manipulation is associated with more quickly subduing protest movements, it may be that other mechanisms are at work. For instance, it may be that election observation makes it harder for front-line agents to avoid personal consequences for engaging in manipulation [@Harvey2022].

By showing that the infrastructure of fraud can actually reduce the cost of post-election protest to the regime, this result poses a challenge to the theory of democratization by election [@Schedler2013;@Morgenbesser2019;@edgellWhenWhereElections2018]. If the capacity to engage in widespread manipulation makes it easier to both tolerate the existence of an opposition and to suppress it during periods of contestation [@dahlPolyarchyParticipationOpposition1971], electoral authoritarianism may be something of an institutional cul-de-sac: the capacity to generate severe election fraud makes it appealing for closed authoritarian regimes to introduce elections, while the manageable risk of protest makes it unappealing to further democratize. Likewise, the results suggest that, on average in authoritarian regimes, the signaling benefits of heavy manipulation outweigh the legitimacy costs incurred by incumbents for such malfeasance [@birch2011]. Future research might investigate the scope conditions for this pattern; cases where the legitimacy costs of fraud are very high may prove to be instances where the 'regime-weakness' model is more appropriate.




<!--


```{r, include=F, warning=FALSE, message=FALSE, echo=FALSE}
## Anova for winning vs. losing
anova_dat <- nelda2.sub %>% dplyr::select(inc.lose, v2elirreg.inv,  jucon.1lag, legcon.1lag, gdpgro.1lag,  osorg.1lag, natlpartyorg.lag,  urban.pct, n.events.pre, state.owned.1lag, patronage.appts.1lag, v2elembaut)

# boxplots and t-tests for the 4 variables at once
for (i in 2:ncol(anova_dat)) { # variables to compare are variables 2 to end
  boxplot(anova_dat[, i] ~ anova_dat$inc.lose, # draw boxplots by group
    ylab = names(anova_dat[i]), # rename y-axis with variable's name
    xlab = "Incumbent loss"
  )
  print(paste(c("ANOVA for variable ", (names(anova_dat[i])))))
  print(t.test(anova_dat[, i] ~ anova_dat$inc.lose)) # print results of t-test
}


```
-->
Finally, this paper does not examine the role of protest risk in insulating established democracies from democratic backsliding [@fearon_self-enforcing_2011]. This question deserves further investigation, but it is possible to draw a few implications from this study. For instance, it may be that the 'regime-strength' model of manipulation is better suited to electoral autocracies, while the 'regime-weakness' model remains applicable in democracies. It is likely that the proportion of elites' political strength derived from electoral legitimacy (rather than from coercive power or patron-client ties) is greater in democracies than in authoritarian regimes. As a result, low popularity may be more closely associated with political weakness in democracies, and electoral manipulation in democracies may be more likely to provoke elite disloyalty than loyalty. At the same time, however, liberal democracies should be constrained against large-scale election manipulation in the first place, as they lack the patronage resources and arbitrary enforcement of laws needed to resolve principal-agent problems at scale. Future research could probe these possible scope conditions. 


<!--
It is also worth noting that mass electoral protest can serve an illiberal purpose, as illustrated by the events of January 6, 2020 in the United States or 2022-2023 in Brazil. Indeed, out of the 156 instances of post-election protest in this dataset, 10 occurred after relatively clean elections (*intentional irregularities* < 0)---a small but not minuscule percentage. These incidents demonstrate that, while crying fraud may sometimes be a useful tool for protest entrepreneurs, losing an election alone can be a catalyst for protest, irrespective of the actual level of election manipulation. This possibility, which would be puzzling if fraud indicates weakness, highlights the often non-strategic nature of protest initiation while also underlining the fact that the course of protest is often driven more by the relative resources and opportunities available to the disaffected and the incumbent government.
-->

These results remain consistent under a several alternate specifications presented in an online appendix. The results are supported using two alternative measures of election manipulation, one that covers a broader array of electoral violations, and another that is drawn directly from election observer reports. Additional tests show that incumbents intensify their efforts to manipulate when protest risk is high, that higher capacity for manipulation makes incumbent election loss less likely, and that the small number of successful protests occur across the spectrum of manipulation. This helps rule out an interpretation of the main results in which weaker regimes choose not to manipulate and thus avoid protest.  The results also remain consistent when the selection criteria are relaxed, and a set of electoral democracies are included in the dataset.  



# Conclusion

There are competing views on the relationship between election fraud and protest. In one family of models, more manipulation indicates greater weakness against post-election challenges. As a result, incumbents may hold cleaner elections when the risk of post-election protest increases. Other models contend that election fraud can indicate incumbent strength. Building on the principal-agent literature of election fraud, I argue that large-scale fraud draws on the same resources that make incumbents resilient to protest, even when regime popularity (and thus vote-shares) are low. Large-scale election manipulation requires incumbents to have control over patronage resources and the ability to protect front-line agents from the consequences of illegal behavior. For elites who are embedded in these patronage networks, extensive electoral manipulation is a credible, intelligible signal of incumbent resources and capacities; these signals help induce elite loyalty during challenges. As a result, I argue, more extensive electoral manipulation should be associated with indicators of regime strength. Resolving the tension between these views in the literature has important implications for the study of election fraud and authoritarian stability.

This paper tests the implications of these two models by investigating the relationship between election fraud, protest risk, and the duration of protest. Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 659 elections in electoral authoritarian regimes to show that 1) pre-election indicators of higher protest risk are associated with more manipulation, not less, and 2) incumbents that engage in larger-scale election fraud are able to more quickly defeat protest movements.

All together, these results suggest that protest risk is unlikely to serve as a constraint on election manipulation in authoritarian regimes. In fact, the indicate the opposite. Incumbents have perverse incentives to manipulate more extensively as risks of post-election contention increases. Those incumbents who successfully produce extensive manipulation---by overcoming large-scale principal-agent problems---defeat post-election challenges more swiftly. In other words, the most effective election manipulators are also the most well-equipped to defeat protest movements. The frailty of protest risk as a guardrail suggests that electoral authoritarianism as a regime-type is a stubborn equilibrium. This is a normatively gloomy picture, and suggests that researchers should further consider other mechanisms beyond mass protest in their models of electoral manipulation.


# Data availability statement

The data analyzed in this study will be made available at the Dataverse upon acceptance of the paper.

\newpage
# References



