---
title: "Appendix part 2"
author: "Cole J. Harvey (Oklahoma State University)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    bookdown::pdf_document2: default
linkcolor: blue
bibliography: C:/Users/colej/Documents/Research projects/bib_ch.bib
#biblio-style: "apalike"
fontsize: 12pt
link-citations: yes
toc: false
csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
header-includes: \usepackage{booktabs}
---


```{r setup, include = FALSE}
library(tidyverse)
library(countrycode)
library(readxl)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
library(stargazer)
library(ggeffects)
library(survival)
library(modelsummary)
library(glmmTMB)
library(huxtable)
library(kableExtra)
library(simPH)


nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2024-02-06.csv"))

nelda2.sub <- nelda2.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
nelda2.sub <- nelda2.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))



##Naming nelda variables

nelda2.sub <- nelda2.sub %>% rename(boycott = nelda14)
nelda2.sub <- nelda2.sub %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
                                                                                 NA)))

##Creating economic crisis variable

nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = as.factor(econ.crisis))

nelda2.sub <- nelda2.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))

nelda2.sub <- nelda2.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))



## Removing cases where inc. loses
#nelda2.sub <- nelda2.sub %>% filter(inc.lose == 0)  #See how this affects things

##Intmon as a factor
nelda2.sub <- nelda2.sub %>% mutate(v2elintmon = as.factor(v2elintmon))


nelda2.sub <- nelda2.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))

#Square of physint

nelda2.sub <- nelda2.sub %>% mutate(physintlag.sq = physinteg.1lag^2)

## Categorical variable for incumbent status

nelda2.sub <- nelda2.sub %>% mutate(incumb.status = as.factor(ifelse(inc.termlimit == "yes" & chosen.successor == 1, "Chosen successor",
                                                           ifelse(inc.termlimit == "yes" & chosen.successor == 0, "Contested exec. election", 
                                                                  ifelse(inc.termlimit == "no", "Executive eligible", "Legislative election"))))) 
nelda2.sub <- nelda2.sub %>% mutate(incumb.status = relevel(incumb.status, ref = "Legislative election"))


#Coding as factors

nelda2.sub <- nelda2.sub %>% mutate(presidential = factor(presidential))
#nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = factor(inc.termlimit.num))
nelda2.sub <- nelda2.sub %>% mutate(rentierstate = factor(rentierstate))
nelda2.sub <- nelda2.sub %>% mutate(taxstate = factor(taxstate))
nelda2.sub <- nelda2.sub %>% mutate(unscheduled_election = factor(unscheduled_election))
nelda2.sub <- nelda2.sub %>% mutate(inc.lose = factor(inc.lose))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))


#Successful protest = new elections or new incumbent

nelda2.sub <- nelda2.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))



# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections

nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))

## Replacing NA with 0 for protest size

#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(is.na(max_size)==T & n.events.post == 0, 0, max_size))

## Fixing missing values of winner.share.upd for liberal democracies

nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(is.na(winner.share.upd)==T & v2x_regime == 3, winner.margin, winner.share.upd))

## Set max_size zero for elections with no protest

#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))

## Excluding liberal democracies

nelda2.sub <- nelda2.sub %>% filter(v2x_regime == 1 | regime.1lag == 1 | v2x_regime == 2 | regime.1lag == 2  ) #Includes only authoritarian, add e-dems with | v2x_regime == 2  #Current version includes lagged 1 in order to capture pre-election auth.

## Remove outlier
nelda2.sub <- nelda2.sub %>% filter(d.events.num <= 100)

#Protest cases only

protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1 & inc.lose == 0)

## Coding median size as 0 when n.events = 0

nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(n.events.post == 0, 0, median.participants))

## Any protest

nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(n.events.post > 0, 1, 0))

## Surv status
library(survival)
library(ggsurvfit)
nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))

```




# Including electoral democracies

The theory articulated in the main text is primarily concerned with dynamics within authoritarian regimes, where patronage resources can be marshaled to support election-manipulation machines, and rule of law can be subverted to protect election-manipulating agents. These mechanisms should be expected to become attenuated, on average, as regimes become more democratic. As such, the main analysis is confined to authoritarian regimes (as coded in V-Dem) where multiparty elections are held. However, this approach imposes a binary cutoff on what is a continuum of regimes, and could introduce some bias by excluding edge cases.

To check this possibility, the analysis here shows the results of models from the main text using a broader range of cases: including both the authoritarian regimes from the original analysis and electoral democracies (V-Dem variable v2x_regime = 2). I expect that changing the sample in this way will weaken the results somewhat, due to the inclusion of cases where theoretically important variables are weaker. For example, the average level of *judical constraints on the executive* in the original sample is 0.37, while the average in the electoral democracies in almost twice as high at 0.66. Similarly, the average incumbent vote-share in the original sample is 0.55, while in the electoral democracies it is 0.41. The liberal democracies have substantially more professionalized bureaucratic appointments on average (0.27) than the electoral authoritarian regimes (-0.33). Consequently I expect levels of electoral manipulation to be lower and less variable in the new sample, on average, if only because the electoral democracies have less developed tools to engage in large-scale fraud. 

Though the results are somewhat weaker, as expected, the general patterns hold. Table \@ref(tab:manipdv-ed) shows how levels of electoral manipulation respond to varying levels of pre-election risk to incumbents. As in the main text, the results show that more pre-election protest is associated with higher levels of fraud. We also see that higher levels of manipulation are seen at middle levels of repression, as in the main text. Two difference are apparent, however. Adding a number of higher-values cases reveals a straightforward, negative, statistically significant relationship between physical integrity protection and election fraud. However, goodness-of-fit measures indicate that the curvilinear relationship fits the data substantially better. Finally, while *regional protest success* has a positive coefficient, it is not statistically significant as it is in the main text. Despite the less theoretically appropriate sample, we still observe the predicted relationship in two models. In a a third, we see the coefficient is in the predicted direction, but loses its statistical significance. This is broadly supportive of the main argument.

```{r manipdv-ed, echo=FALSE}

m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag +  v2elembaut + (1 | stateid), data = nelda2.sub) 
#summary(m1a)

m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag  + unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub) 

#summary(m1b)

m1c <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag  + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub) 
#summary(m1c)

m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag + patronage.appts.1lag + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub) 

#summary(m1)

m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag + v2elembaut +  (1 | stateid), data = nelda2.sub) 
#summary(m1a)

m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag +  unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub) 

#summary(m1b)

models <- list()
models[['Model 16']] <- m1
models[['Model 17']] <- m1b
models[['Model 18']] <- m1c
models[['Model 19']] <- m1a


modelsummary(models, output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Random-intercept multilevel model of election-day fraud', coef_omit = "SD", coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity",
                  "state.owned.1lag" = "Private ownership (lag)",
                 "patronage.appts.1lag" = "Professional bureaucracy (lag)",
                 "v2elembaut" = "Election management body autonomy"
                 )) %>% kable_styling(latex_options = "HOLD_position")

##Note: state owned variable is inverse coded, so should be labeled as private ownership in economy
##Note: Patronage appointments ialso inverse coded. Higher values = more professionalism
```



Turning to models of protest initiation, the findings from the main text are upheld. Higher levels of fraud are associated with a higher risk of protest, but only when incumbent vote-shares are low. Table \@ref(tab:logit-table-edem) reports these results, and Figure \@ref(fig:margin-prob-fig-edem) shows the predicted probabilities.


```{r include=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))


model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
summary(model.margin.binary)


model.margin.binary.int <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
summary(model.margin.binary.int)


model.margin.binary_noc <- glm(any_protest ~ jucon.1lag + legcon.1lag +  altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
summary(model.margin.binary_noc)



## Plots

####Margin

#plot_binary_margin_max <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-2.3, 3.2]"))
#plot_binary_margin_max <- plot_binary_margin_max +  labs(x = "", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()

plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c( "winner.share.upd", "v2elirreg.inv [-.05, 2]"), show.data = T)
plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()


```

```{r logit-table-edem, echo=FALSE}
models2 <- list()
models2[['Model 20']] <- model.margin.binary
models2[['Model 21']] <- model.margin.binary.int
models2[['Model 22']] <- model.margin.binary_noc

modelsummary(models2, output = "kableExtra",  stars = T, title = 'Logit models of protest occurrence', coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity",
                 "v2elirreg.inv" = "Election fraud"
                 )) %>% kable_styling(latex_options = "HOLD_position")
```

```{r margin-prob-fig-edem, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
plot_binary_margin_typ_n
```

Lastly, the survival models also support the findings of the main text, as Table \@ref(tab:cox-models-tab-edem) and Figure \@ref(fig:margin-dur-fig-edem) show. Higher levels of election fraud are associated with shorter-lived protest movements when incumbent vote-shares are low, even in the expanded sample. 

```{r, include=FALSE}
library(survival)
library(ggsurvfit)
nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
nelda2.sub <- nelda2.sub %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
                                                                       ifelse(d.events.num >= 69, 0, 1)))








cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
summary(cox1)




cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
summary(cox2)
p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.73, 1.79]")) + coord_cartesian(ylim = c(0, 10)) #This is the 10th and 90th quantiles
p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ration", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw() #These are extreme values: 15 and 85th percentiles



```


```{r cox-models-tab-edem, echo=FALSE, warning=FALSE}
models3 <- list()
models3[['Model 23']] <- cox1
models3[['Model 24']] <- cox2


modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Cox proportional hazard models of protest duration', coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", 
                 "winner.share.upd" = "Incumbent vote-share",
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "v2elirreg.inv" = "Election fraud",
                 "n.events.post" = "N. post-election protests"
                 ))

```

```{r margin-dur-fig-edem, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
p.duration.margin
```






\newpage

# References

